---
title: "Camera Calibration Workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Camera Calibration Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.align = "center"
)
library(radiatR)
```

Tracks recorded from an overhead camera live in *pixel* coordinates that are
distorted by the lens and offset from the arena centre. To analyse movement in a
true metric (or at least undistorted) frame you need the camera's calibration:
its intrinsic matrix and distortion coefficients.

**radiatR does not estimate calibrations itself.** Estimating intrinsics from a
checkerboard sequence is a solved problem with mature, well-tested
implementations — MATLAB's *Computer Vision Toolbox*, OpenCV's
`calibrateCamera`, and others. radiatR's job is to *import* the coefficients
those tools produce and *apply* them to your tracks. This vignette covers that
import-and-correct workflow.

## The model radiatR applies

radiatR uses the pinhole + Brown-Conrady model shared by MATLAB and OpenCV:

- **Focal lengths** `fx`, `fy` and **principal point** `cx`, `cy`, in pixels.
- **Radial distortion** `k1`, `k2`, `k3` and **tangential distortion** `p1`,
  `p2`.

These are stored in a [`CalModel`](../reference/CalModel-class.html) object,
which `calibrate_positions()` applies to a `TrajSet`.

A subtlety worth flagging up front: internally radiatR keeps the intrinsic
matrix in MATLAB's *transposed* convention — focal lengths on the diagonal,
principal point in the **bottom row** (`K[3, 1:2]`), not the right-hand column.
You never have to assemble that matrix by hand: `cal_model()` and
`read_calibration()` own it for you.

## Building a model from known coefficients

If you already have the numbers — from a calibration report, a paper, or a
colleague — `cal_model()` assembles the `CalModel` directly:

```{r cal-model}
model <- cal_model(
  fx = 800, fy = 800,   # focal lengths (px)
  cx = 640, cy = 360,   # principal point (px)
  k1 = -0.28, k2 = 0.05,
  p1 = 1e-3,  p2 = -5e-4
)
model
```

### What `F` controls

The `F` argument sets the units of the corrected output. When `F` is left as
`NULL` (the default) it is set to `c(fx, fy)`, which means the output stays in
**undistorted pixels** — distortion and the principal-point offset are removed,
but the scale is unchanged. To obtain a metric result instead, pass the physical
focal length in millimetres:

```{r cal-model-mm}
model_mm <- cal_model(fx = 800, fy = 800, cx = 640, cy = 360,
                      k1 = -0.28, k2 = 0.05, F = 20)
```

## Importing a calibration file

In practice you will export the calibration from another tool and read it with
`read_calibration()`, which auto-detects the format from the file extension:

```{r read-calibration, eval = FALSE}
# MATLAB Computer Vision Toolbox (.mat)
model <- read_calibration("camera_params.mat")

# OpenCV FileStorage (.yml / .yaml / .json)
model <- read_calibration("opencv_calib.yml")

# Plain CSV (wide or parameter/value)
model <- read_calibration("calib.csv")
```

### MATLAB

Save the relevant fields of a `cameraParameters` / `cameraIntrinsics` object to
a `.mat` file. radiatR recognises both the pre-R2022b `IntrinsicMatrix` (already
in the transposed convention) and the newer standard `K`, alongside
`RadialDistortion` and `TangentialDistortion`. Reading `.mat` files needs the
**R.matlab** package.

```matlab
% in MATLAB, after calibrating
params = cameraParams;
IntrinsicMatrix     = params.IntrinsicMatrix;
RadialDistortion    = params.RadialDistortion;
TangentialDistortion = params.TangentialDistortion;
save('camera_params.mat', 'IntrinsicMatrix', ...
     'RadialDistortion', 'TangentialDistortion');
```

MATLAB reports the principal point in **1-based** pixels. `read_calibration()`
subtracts one by default (`principal_base = 1`) so it lines up with the 0-based
coordinates produced by most tracking tools. Override `principal_base` if your
tracker uses a different convention.

### OpenCV

`cv::FileStorage` writes a `camera_matrix` and `distortion_coefficients` (in the
order `k1, k2, p1, p2, k3`). radiatR reads both the YAML and JSON variants and
tolerates the non-standard `%YAML:1.0` header and `!!opencv-matrix` tags. YAML
needs the **yaml** package; JSON needs **jsonlite**. OpenCV principal points are
already 0-based, so no shift is applied by default.

### Plain CSV

For coefficients typed out by hand, a CSV works in either layout — wide:

```
fx,fy,cx,cy,k1,k2,k3,p1,p2
800,800,640,360,-0.28,0.05,0,0.001,-0.0005
```

or long (`parameter,value` rows). Missing coefficients default to zero. CSV
principal points are treated as 0-based.

## Correcting points

For a single landmark, `cam_cal_pt()` applies the model to one pixel coordinate;
`cam_cal_many()` does the same for a table of points in one pass:

```{r points}
pts <- matrix(
  c(640, 360,
    700, 420,
    580, 320),
  ncol = 2, byrow = TRUE,
  dimnames = list(NULL, c("x", "y"))
)
cam_cal_many(pts, model@K, model@k, model@F)
```

The point at the principal point `(640, 360)` maps to the origin; others are
recentred, undistorted, and scaled by `F`.

## Correcting a whole TrajSet

The usual entry point is `calibrate_positions()`, which corrects every track in
a `TrajSet`, rewrites its `x`/`y` columns in place, and re-derives `angle` from
the corrected geometry. The bundled millipede example provides real pixel-space
tracks from a camera above a circular arena.

```{r trajset-example}
track_dir <- system.file("extdata", "tracks", package = "radiatR")
manifest_path <- system.file("extdata", "millipede_trials.csv", package = "radiatR")

file_tbl <- import_tracks(track_dir)
manifest <- import_info(manifest_path)
file_tbl <- load_tracks(file_tbl, manifest, track_dir)

# Read the first track file in pixel space
track_raw <- read.delim(
  file.path(track_dir, file_tbl$track[1]),
  sep = "\t", header = FALSE
)[, 1:3]
names(track_raw) <- c("frame", "x", "y")
track_raw$id <- tools::file_path_sans_ext(basename(file_tbl$track[1]))

ts_px <- TrajSet(
  track_raw,
  id = "id",
  time = "frame", x = "x", y = "y",
  normalize_xy = FALSE
)

ts_mm <- calibrate_positions(ts_px, model)
head(ts_mm@data)
```

In a real workflow, replace the illustrative `model` above with one imported via
`read_calibration()` from a calibration captured with the same camera and lens.

## Convergence diagnostics

Removing distortion requires inverting the distortion model, which `radiatR`
does with a fixed-point solver (as OpenCV's `undistortPoints` does). The defaults
(`max_iterations = 5`, `tolerance = 1e-9`) are usually ample. For extreme
distortion or points far from the optical axis you can relax them:

```{r convergence, eval = FALSE}
cam_cal_pt(650, 380, model@K, model@k, model@F,
           max_iterations = 10, tolerance = 1e-12)
```

If convergence fails the function returns the last iterate; check your
coefficients or tighten the stopping criteria.

## Summary

- radiatR imports calibrations; it does not estimate them. Calibrate in MATLAB,
  OpenCV, or another tool, then bring the coefficients in.
- `read_calibration()` reads MATLAB `.mat`, OpenCV YAML/JSON, and plain CSV;
  `cal_model()` builds a model from coefficients you already hold.
- Apply the resulting `CalModel` to single points (`cam_cal_pt()`), tables
  (`cam_cal_many()`), or whole trajectory sets (`calibrate_positions()`).
- Mind the conventions: principal-point pixel base (1-based MATLAB vs 0-based
  OpenCV) and the `F` scale factor (undistorted pixels by default, metric if you
  pass a physical focal length).
