---
title: "radiatR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{radiatR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
if (requireNamespace("pkgload", quietly = TRUE)) {
  # Always load from source so the vignette reflects the current development state.
  pkgload::load_all("..", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else if (requireNamespace("radiatR", quietly = TRUE)) {
  library(radiatR)
} else {
  stop("Package 'radiatR' not installed and 'pkgload' not available.")
}
library(ggplot2)
```

## Overview

`radiatR` streamlines the journey from headings and trajectories in circular space to publication-ready radial plots. Angles can be supplied directly, or reconstructed from movement trajectories. The package includes:

-   utilities for importing paired landmark/track text files produced by popular tracking suites;
-   helpers for extracting trial limits and computing circular summary statistics;
-   polished `ggplot2` layers for drawing concentric guides and annotated paths.

## Import Pipeline

The package ships a full set of example tracks (a millipede orientation experiment) that demonstrate the import workflow. Landmark files (`_point01.txt`) record two pixel-space reference points per trial (the origin and a target location on the circumference); track files (`_point02.txt`) contain the full xy trajectory.

```{r load-demo}
track_dir     <- system.file("extdata", "tracks", package = "radiatR")
manifest_path <- system.file("extdata", "millipede_trials.csv", package = "radiatR")

file_tbl  <- import_tracks(track_dir)
manifest  <- import_info(manifest_path)
file_tbl  <- load_tracks(file_tbl, manifest, track_dir)

ts_demo <- suppressWarnings(get_all_object_pos(file_tbl = file_tbl, track_dir = track_dir))
ts_demo
```

`get_all_object_pos()` reads each landmark/track pair, normalises coordinates to a unit circle (arena radius = 1), and returns a `TrajSet`. Trial metadata (arena radius, target position, frame limits) is in `ts_demo@meta$trial_limits`.

## Full Millipede Dataset

The package also provides `cpunctatus`, a pre-computed `TrajSet` of all 235 *Cylindroiulus punctatus* trajectories across the stimulus conditions (target half-widths of 5, 10, 15, 20, 30, 40, 50 degrees, plus a featureless control with `arc = 0`). Loading it is instant, and the per-trial target half-width is already attached as the `arc` column.

```{r load-full}
data(cpunctatus)
cpunctatus
```

## Plotting Trajectories

`radiate()` draws trajectories on a unit circle with concentric reference rings. Colouring by the `arc` factor shows how paths cluster by condition.

```{r plot-all}
radiate(cpunctatus,
        group_col  = "trial_id",
        colour_col = "arc",
        show_labels = FALSE,
        show_arrow  = FALSE,
        display    = circ_display(zero = 0))
```

## Heading Overlays

The **crossing method** — projecting the vector between two concentric-ring crossings to the unit circle — assigns one directional heading per trial. `derive_headings()` with `return_coords = TRUE` returns both the heading angle and the inner-ring crossing position.

```{r headings}
hd <- derive_headings(cpunctatus, rule = "crossing",
                      circ0 = 0.2, circ1 = 0.4,
                      return_coords = TRUE)
names(hd)[names(hd) == "id"] <- "trial_id"
# join the target half-width (arc) from the dataset for grouping/faceting
arc_map <- unique(cpunctatus@data[, c("trial_id", "arc")])
hd       <- merge(hd, arc_map, by = "trial_id")
hd$arc   <- factor(hd$arc)
attr(hd, "colour_col") <- "arc"
attr(hd, "display")    <- circ_display(zero = 0)
head(hd[, c("trial_id", "arc", "heading", "x_inner", "y_inner")])
```

Overlaying the heading endpoints (one hollow circle per trial) and the grand mean direction on the combined trajectory plot gives a compact summary of the full dataset:

```{r headings-overlay}
p_all <- radiate(cpunctatus,
                 group_col   = "trial_id",
                 colour_col  = "arc",
                 show_labels = FALSE,
                 show_arrow  = FALSE,
                 display     = circ_display(zero = 0)) +
  add_heading_points(hd, colour_col = "arc", size = 1, alpha = 0.6)

p_all + add_heading_arrow(hd)
```

```{r summ-display, include=FALSE}
summ_hd <- compute_circ_mean(hd)
# Convert UC mean_dir to display degrees: circ_display(zero=0, clockwise=TRUE)
# => display_deg = (-uc_angle %% 2pi) * 180/pi
summ_hd_deg <- round(((-summ_hd$mean_dir) %% (2 * pi)) * 180 / pi, 1)
summ_hd_R   <- round(summ_hd$resultant_R, 2)
```

The grand mean arrow points at `r summ_hd_deg`° relative to the reference direction (clock convention; 0° = toward reference) with *R* = `r summ_hd_R`, reflecting the overall reference-relative tendency across all conditions.

## Circular Interval Arc

Three functions handle directional uncertainty arcs in parallel with the density overlay functions:

| Function | Role |
|---|---|
| `compute_circ_interval()` | Computes arc bounds from raw headings; returns a data frame with `lower`, `upper`, `mean_dir`, and `wraps` |
| `add_circ_interval()` | Renders any bounds data frame as an arc at a configurable radius — agnostic to how the bounds were produced |
| `add_heading_interval()` | Convenience wrapper: calls the two above in sequence |

Two statistics are supported: `stat = "bootstrap_ci"` bootstraps the von Mises MLE confidence interval for the mean direction; `stat = "sd"` draws a ±1 circular SD arc. The split design lets you substitute Bayesian credible bounds into the data frame before rendering:

```{r interval-concept, eval=FALSE}
iv <- compute_circ_interval(hd, colour_col = "arc", stat = "bootstrap_ci")
# replace with Bayesian posteriors: iv$lower <- ...; iv$upper <- ...
add_circ_interval(iv, colour_col = "arc")
```

## Building Up a Panel

Layers compose with the standard `ggplot2` `+` operator, so a plot can be assembled feature by feature. The four chunks below start from trajectories only and add heading endpoints, the grand mean arrow, and finally a bootstrap CI arc.

**Trajectories:**

```{r build-1}
p <- radiate(cpunctatus,
             group_col   = "trial_id",
             colour_col  = "arc",
             show_labels = FALSE,
             show_arrow  = FALSE,
             display     = circ_display(zero = 0))
p
```

**+ Heading endpoints** at each trial's crossing location:

```{r build-2}
p <- p + add_heading_points(hd, colour_col = "arc", size = 1.5, alpha = 0.7)
p
```

**+ Grand mean direction arrow:**

```{r build-3}
p <- p + add_heading_arrow(hd)
p
```



The arc at radius 1.05 spans the 95 % bootstrap confidence interval for the grand mean direction pooled across all arc conditions.

## Colour Options

Three strategies control how trajectories are coloured.

**Option 1 — single colour.** Pass a colour string directly to the heading-overlay helpers; omit `colour_col` from `radiate()` to draw all tracks in the default grey.

**Option 2 — cycling palette.** `colour_cycle` assigns each trajectory a colour index that cycles back to 1 after every *n* trajectories. When `panel_by` is set the cycle restarts independently within each panel so that trajectory 1 in every panel always gets colour 1. Pass an integer for automatic palette assignment, or a character vector to specify colours explicitly.

**Option 3 — variable mapping.** `colour_col` maps any data column to colour (e.g. the `arc` factor used in the combined-trajectory plot above).

## Per-Condition Panels

Faceting by arc angle shows each condition separately. Within each panel, `assign_cycle_colours()` distinguishes individual trajectories by cycling through 10 colours, resetting at each panel boundary. Calling it explicitly on both the track data and the headings data frame (joining by trial id) means heading markers inherit the exact per-trajectory colour — not a single per-condition colour — when `colour_col` is used for both.

```{r arc-panels-colours}
# Pre-compute cycling colours: 10 colours, restarting within each arc panel.
# Join to headings so heading layers can use the same column.
cpunctatus_cc        <- cpunctatus
cpunctatus_cc@data   <- assign_cycle_colours(cpunctatus@data,
                                            id_col    = "trial_id",
                                            n         = 10,
                                            panel_col = "arc")
colour_map <- unique(cpunctatus_cc@data[, c("trial_id", "cycle_colour")])
hd_cc      <- merge(hd, colour_map,
                    by      = "trial_id", all.x = TRUE)
attr(hd_cc, "display")    <- attr(hd, "display", exact = TRUE)
attr(hd_cc, "colour_col") <- "cycle_colour"
```

```{r arc-panels-simple, fig.width=8, fig.height=5}
p <- radiate(cpunctatus_cc,
        group_col    = "trial_id",
        colour_col   = "cycle_colour",
        panel_by     = "arc",
        ncol         = 3,
        show_labels  = FALSE,
        show_arrow   = FALSE,
        display      = circ_display(zero = 0))
p
```

**+ Bootstrap CI arc** added first so it sits behind heading markers:

```{r build-4, fig.width=8, fig.height=5}
p <- p + add_heading_interval(hd_cc, colour_col = "arc", colour = "black",
                               stat = "bootstrap_ci", boot_reps = 999L)
p
```

**+ Heading vectors** (dotted lines from inner crossing to perimeter):

```{r arc-panels-vectors, fig.width=8, fig.height=5}
p <- p + add_heading_vectors(hd_cc)
p
```

**+ Heading points** on top of the CI arc:

```{r arc-panels-headings, fig.width=8, fig.height=5}
p <- p + add_heading_points(hd_cc, size = 4)
p
```

```{r arc-panels, fig.width=8, fig.height=5}
p <- p + add_heading_arrow(hd_cc, colour_col = "arc", colour = "black")
p
```

Each panel shows trajectories and heading markers coloured by per-trajectory cycling palette, a bootstrap CI arc at radius 1.05 in the condition colour (rendered behind the heading points), a solid mean direction arrow, and degree labels. All in clock convention (0° = toward reference, clockwise). Use `strip_position = "inside"` to place the label inside the plot area, or `strip_labels = FALSE` to suppress it.

## Circular Density Overlays

Three functions handle directional density overlays; they are designed to compose cleanly with any density source:

| Function | Role |
|---|---|
| `compute_circular_density()` | Estimates density from raw headings; returns a plain data frame of `(theta, density)` pairs |
| `add_circular_density()` | Renders any `(theta, density)` data frame as a radial overlay — agnostic to how the density was produced |
| `add_heading_density()` | Convenience wrapper: calls the two above in sequence |

Three built-in methods are available in `compute_circular_density()`: `"vonmises"` (MLE via `circular::mle.vonmises()`), `"kernel"` (circular KDE), and `"histogram"` (bin counts). Because the computation and rendering steps are separate, the `density` column can be replaced with values from any other model — including a Bayesian posterior predictive density from `brms` — before calling `add_circular_density()`:

```{r density-concept, eval=FALSE}
dens_df <- compute_circular_density(hd, colour_col = "arc", method = "vonmises")
# replace with Bayesian posterior: dens_df$density <- my_brms_fitted_density(dens_df$theta)
add_circular_density(dens_df, colour_col = "arc", fill = "grey80", alpha = 0.35)
```

The convenience wrapper `add_heading_density()` skips the intermediate step when no substitution is needed. The combined panel plot below uses it to shade a per-condition von Mises density alongside the heading vectors:

```{r density-overlay, fig.width=8, fig.height=5}
radiate(cpunctatus_cc,
        group_col    = "trial_id",
        colour_col   = "cycle_colour",
        panel_by     = "arc",
        ncol         = 3,
        show_labels  = FALSE,
        show_arrow   = FALSE,
        display      = circ_display(zero = 0)) +
  add_heading_density(hd_cc, colour_col = "arc",
                      method = "vonmises", scale = 0.4,
                      fill = "grey80", alpha = 0.35) +
  add_heading_vectors(hd_cc) +
  add_heading_arrow(hd_cc, colour_col = "arc", colour = "black")
```

The shaded region is the von Mises density fitted to the crossing headings within each arc condition. A narrower, taller peak indicates more concentrated directional responses. Switch to `method = "kernel"` for a non-parametric estimate, or `method = "histogram"` for a raw count display.

## Bootstrap Confidence Band

`compute_circular_density()` with `boot_reps > 0` runs a non-parametric bootstrap: heading samples are drawn with replacement, a von Mises MLE is fitted to each replicate, and the density is re-evaluated on the same angular grid. The `boot_alpha / 2` and `1 - boot_alpha / 2` quantiles across replicates become `density_lower` and `density_upper` columns, which `add_circular_density()` renders as a shaded band around the fitted curve.

```{r bootstrap-ci, fig.width=8, fig.height=5}
dens_boot <- compute_circular_density(hd, colour_col = "arc",
                                      method = "vonmises",
                                      boot_reps = 499L, boot_alpha = 0.05,
                                      n_theta = 300L)

radiate(cpunctatus,
        group_col    = "trial_id",
        colour_cycle = 10,
        panel_by     = "arc",
        ncol         = 3,
        show_labels  = FALSE,
        show_arrow   = FALSE,
        display      = circ_display(zero = 0)) +
  add_circular_density(dens_boot, colour_col = "arc",
                       scale = 0.4, fill = "grey80", alpha = 0.35,
                       ci_fill = "grey60", ci_alpha = 0.35) +
  add_heading_arrow(hd, colour_col = "arc", colour = "black")
```

The darker band is the 95% bootstrap confidence interval for the fitted von Mises density. A wider band reflects greater parametric uncertainty, typically seen in the smaller or more diffuse conditions. The `density_lower` and `density_upper` columns in `dens_boot` can be replaced with interval values from a Bayesian model (e.g. the 2.5th and 97.5th percentiles of a `brms` posterior predictive distribution) before passing to `add_circular_density()`.

## Circular Summary Statistics

`compute_circ_mean()` returns the per-condition statistics behind the arrows above:

```{r stats}
compute_circ_mean(hd, colour_col = "arc")[, c("arc", "mean_dir", "resultant_R")]
```

For within-trial path consistency (tortuosity), `circ_summary()` operates on the step-by-step angle distribution:

```{r within-trial}
circ_summary(ts_demo)
```

High resultant lengths (close to 1) indicate very consistent step directions within a trial.

## Alternative Heading Rules

`derive_headings()` supports fourteen built-in rules. `"crossing"` is well
suited to echinoderm-style tracks — moderate tortuosity, consistent outward
movement — but other rules may be more appropriate depending on the taxon and
experimental design. Use `list_heading_rules()` to see all available names;
custom rules can be added with `register_heading_rule()`.

Two parameter-free alternatives are especially useful:

| Rule | What it returns | Typical use |
|---|---|---|
| `"distal"` | Angular position of the frame with the largest radius | Straight outward paths (dung beetles, ballistic homing) |
| `"net"` | Direction of the start-to-end displacement vector | Sinuous or multi-phase paths where only net displacement matters |

### distal

The **distal** rule takes `atan2(y, x)` at the frame where the animal is
farthest from the arena centre. It requires no ring parameters and never
returns `NA` because every trial has a most-distal frame.

```{r hd-distal}
hd_distal <- derive_headings(cpunctatus, rule = "distal",
                              return_coords = TRUE)
names(hd_distal)[names(hd_distal) == "id"] <- "trial_id"
# join the target half-width (arc) from the dataset, as for the crossing rule
hd_distal     <- merge(hd_distal, arc_map, by = "trial_id")
hd_distal$arc <- factor(hd_distal$arc)
attr(hd_distal, "display") <- circ_display(zero = 0)
```

Per-condition resultant lengths from `"distal"` are very close to those from
`"crossing"`, consistent with millipede tracks being fairly direct:

```{r hd-compare-R}
sm_cross  <- compute_circ_mean(hd,       colour_col = "arc")[, c("arc", "resultant_R")]
sm_distal <- compute_circ_mean(hd_distal, colour_col = "arc")[, c("arc", "resultant_R")]
names(sm_cross)[2]  <- "crossing"
names(sm_distal)[2] <- "distal"
merge(sm_cross, sm_distal, by = "arc")
```

Visualising the distal headings on the per-condition panel confirms that the
spatial pattern is consistent with the crossing method:

```{r hd-distal-panel, fig.width=8, fig.height=5}
cpunctatus_cc_d        <- cpunctatus
cpunctatus_cc_d@data   <- assign_cycle_colours(cpunctatus@data,
                                              id_col    = "trial_id",
                                              n         = 10,
                                              panel_col = "arc")
colour_map_d <- unique(cpunctatus_cc_d@data[, c("trial_id", "cycle_colour")])
hd_distal_cc <- merge(hd_distal, colour_map_d, by = "trial_id", all.x = TRUE)
attr(hd_distal_cc, "display")    <- attr(hd_distal, "display", exact = TRUE)
attr(hd_distal_cc, "colour_col") <- "cycle_colour"

radiate(cpunctatus_cc_d,
        group_col   = "trial_id",
        colour_col  = "cycle_colour",
        panel_by    = "arc",
        ncol        = 3,
        show_labels = FALSE,
        show_arrow  = FALSE,
        display     = circ_display(zero = 0)) +
  add_heading_interval(hd_distal_cc, colour_col = "arc", colour = "black",
                       stat = "bootstrap_ci", boot_reps = 999L) +
  add_heading_points(hd_distal_cc, size = 4) +
  add_heading_arrow(hd_distal_cc, colour_col = "arc", colour = "black")
```

Because millipede tracks are nearly radial, the heading estimates agree
closely across methods. For more sinuous species (moths, pigeons) the choice
of rule has a larger influence on the result.

### net

The **net** rule returns the direction of the vector from the first recorded
position to the last, ignoring all intermediate points. It is the fastest
possible estimate and is the natural match for GPS vanishing-bearing studies
where only the release and final observed positions are available.

```{r hd-net}
hd_net <- derive_headings(cpunctatus, rule = "net")
names(hd_net)[names(hd_net) == "id"] <- "trial_id"
# join the target half-width (arc) from the dataset, as for the crossing rule
hd_net     <- merge(hd_net, arc_map, by = "trial_id")
hd_net$arc <- factor(hd_net$arc)

sm_net <- compute_circ_mean(hd_net, colour_col = "arc")[, c("arc", "resultant_R")]
names(sm_net)[2] <- "net"
Reduce(function(a, b) merge(a, b, by = "arc"),
       list(sm_cross, sm_distal, sm_net))
```

For these tracks the three methods give very similar resultant lengths. Larger
differences would be expected for species with complex search behaviour
before committing to a direction.

## Simulating Demo Data

`simulate_tracks()` generates synthetic trajectories without any files, useful for testing and demonstrations. The `kappa` parameter controls directedness:

```{r simulate}
sim_df <- simulate_tracks(
  conditions = data.frame(n_trials = c(5L, 5L), kappa = c(2, 8)),
  n_points = 150,
  seed = 42
)
ts_sim <- TrajSet(
  sim_df,
  id = "trial_id", time = "frame",
  angle = "rel_theta", x = "rel_x", y = "rel_y",
  angle_unit = "radians", normalize_xy = FALSE
)
radiate(ts_sim,
        group_col    = "trial_id",
        colour_cycle = 5,
        panel_by     = "condition",
        ncol         = 2,
        show_arrow   = TRUE)
```

The left panel (low `kappa`) shows tortuous paths; the right (high `kappa`) shows straighter, more directed tracks. The mean resultant arrow is computed independently per panel.
