---
title: "radiatR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{radiatR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
if (requireNamespace("radiatR", quietly = TRUE)) {
  library(radiatR)
} else if (requireNamespace("pkgload", quietly = TRUE)) {
  # Load package from source when knitting the vignette interactively
  pkgload::load_all("..", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else {
  stop("Package 'radiatR' not installed and 'pkgload' not available. Install with remotes::install_local('.') or install.packages('pkgload').")
}
library(ggplot2)
```

## Overview

`radiatR` streamlines the journey from raw tracking files to publication-ready plots for experiments conducted in circular arenas. The package includes:

-   utilities for importing paired landmark/track text files produced by popular tracking suites;
-   helpers for extracting trial limits and computing circular summary statistics;
-   polished `ggplot2` layers for drawing concentric guides and annotated paths.

## Example Data: *P. lividus* Visual Acuity Trials

The package ships five baseline trials from a *Paracentrotus lividus* (purple sea urchin) visual acuity experiment. Each urchin was placed at the centre of a circular arena (~2 m diameter) and allowed to walk freely toward a patterned wall stimulus (arc = 0°, marble obstacle). Landmark files (`_point01.txt`) record two pixel-space reference points per trial (arena centre and stimulus edge); track files (`_point02.txt`) contain the full xy trajectory.

```{r load-real}
track_dir     <- system.file("extdata", "tracks", package = "radiatR")
manifest_path <- system.file("extdata", "P_lividus_trials.csv", package = "radiatR")

file_tbl  <- import_tracks(track_dir)
manifest  <- import_info(manifest_path)
file_tbl  <- load_tracks(file_tbl, manifest, track_dir)

ts <- get_all_object_pos(file_tbl = file_tbl, track_dir = track_dir)
ts
```

`get_all_object_pos()` reads each landmark/track pair, normalises coordinates to a unit circle (arena radius = 1), and returns a single `TrajSet` containing all trials. Trial metadata is available in `ts@meta$trial_limits`.

```{r plot-real}
radiate(ts, group_col = "trial_id", show_arrow = TRUE)
```

Each arc represents one urchin's path from the arena centre outward. The arrow indicates where the animal first crossed the outer 20 % annulus.

## Circular Summary Statistics

`circ_summary()` computes mean angle, resultant length, and related statistics for a `TrajSet`:

```{r stats}
circ_summary(ts)
```

## Simulating Demo Data

When you need a reproducible toy dataset without any files on disk, `simulate_tracks()` creates tidy data with controllable tortuosity:

```{r simulate}
sim_df <- simulate_tracks(conditions = data.frame(n_trials = 4L), n_points = 150, seed = 42)
ts_sim <- TrajSet(
  sim_df,
  id = "trial_id",
  time = "frame",
  angle = "rel_theta",
  x = "rel_x",
  y = "rel_y",
  angle_unit = "radians",
  normalize_xy = FALSE
)
radiate(ts_sim, group_col = "trial_id", colour_col = "tortuosity")
```

## Advanced: Accessing Underlying Data

```{r trajset-advanced}
head(ts@data)
head(ts@meta$trial_limits)
```
