---
title: "radiatR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{radiatR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
if (requireNamespace("pkgload", quietly = TRUE)) {
  # Always load from source so the vignette reflects the current development state.
  pkgload::load_all("..", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else if (requireNamespace("radiatR", quietly = TRUE)) {
  library(radiatR)
} else {
  stop("Package 'radiatR' not installed and 'pkgload' not available. Install with remotes::install_local('.') or install.packages('pkgload').")
}
library(ggplot2)
```

## Overview

`radiatR` streamlines the journey from raw tracking files to publication-ready plots for experiments conducted in circular arenas. The package includes:

-   utilities for importing paired landmark/track text files produced by popular tracking suites;
-   helpers for extracting trial limits and computing circular summary statistics;
-   polished `ggplot2` layers for drawing concentric guides and annotated paths.

## Example Data: *P. lividus* Visual Acuity Trials

The package ships five baseline trials from a *Paracentrotus lividus* (purple sea urchin) visual acuity experiment. Each urchin was placed at the centre of a circular arena (~2 m diameter) and allowed to walk freely toward a patterned wall stimulus (arc = 0°, marble obstacle). Landmark files (`_point01.txt`) record two pixel-space reference points per trial (arena centre and stimulus edge); track files (`_point02.txt`) contain the full xy trajectory.

```{r load-real}
track_dir     <- system.file("extdata", "tracks", package = "radiatR")
manifest_path <- system.file("extdata", "P_lividus_trials.csv", package = "radiatR")

file_tbl  <- import_tracks(track_dir)
manifest  <- import_info(manifest_path)
file_tbl  <- load_tracks(file_tbl, manifest, track_dir)

ts <- get_all_object_pos(file_tbl = file_tbl, track_dir = track_dir)
ts
```

`get_all_object_pos()` reads each landmark/track pair, normalises coordinates to a unit circle (arena radius = 1), and returns a single `TrajSet` containing all trials. Trial metadata is available in `ts@meta$trial_limits`.

```{r plot-real}
radiate(ts, group_col = "trial_id", colour_col = "trial_id", show_arrow = TRUE)
```

Each arc represents one urchin's path from the arena centre outward. The arrow shows the grand mean direction weighted by resultant length.

## Heading Overlays

The **crossing method** — projecting the vector between two concentric-ring crossings to the unit circle — is one way to assign a single heading per trial. `derive_headings()` with `rule = "crossing"` and `return_coords = TRUE` returns both the heading angle and the inner-ring crossing position needed to draw the extrapolated vector.

```{r headings}
hd <- derive_headings(ts, rule = "crossing",
                      circ0 = 0.2, circ1 = 0.4,
                      return_coords = TRUE)
# align id column name with TrajSet so colour scales merge
names(hd)[names(hd) == "id"] <- "trial_id"
hd[, c("trial_id", "heading", "x_inner", "y_inner")]
```

Layer the heading markers onto the track plot with `+`:

```{r headings-plot}
radiate(ts, group_col = "trial_id", colour_col = "trial_id", show_arrow = FALSE) +
  add_heading_vectors(hd, colour_col = "trial_id") +
  add_heading_points(hd, colour_col = "trial_id", size = 3)
```

Hollow circles mark where each heading vector intersects the unit circle; dotted lines trace back to the inner-ring crossing point. The `show_arrow = FALSE` argument suppresses the grand-mean arrow so the per-trial endpoints are unobscured.

## Circular Summary Statistics

`circ_summary()` computes mean angle, resultant length, and related statistics for a `TrajSet`:

```{r stats}
circ_summary(ts)
```

## Simulating Demo Data

When you need a reproducible toy dataset without any files on disk, `simulate_tracks()` creates tidy data with controllable tortuosity. Pass a `conditions` data frame to generate multiple groups — for example, two levels of directedness controlled by the `kappa` parameter:

```{r simulate}
sim_df <- simulate_tracks(
  conditions = data.frame(n_trials = c(3L, 3L), kappa = c(2, 8)),
  n_points = 150,
  seed = 42
)
ts_sim <- TrajSet(
  sim_df,
  id = "trial_id",
  time = "frame",
  angle = "rel_theta",
  x = "rel_x",
  y = "rel_y",
  angle_unit = "radians",
  normalize_xy = FALSE
)
```

Pass `panel_by` to `radiate()` to split the plot into one panel per level of a grouping variable — analogous to `facet_wrap()` in base ggplot2:

```{r simulate-facet}
radiate(ts_sim,
        group_col  = "trial_id",
        colour_col = "trial_id",
        panel_by   = "condition",
        ncol       = 2,
        show_arrow = TRUE)
```

The left panel (`condition_1`, low `kappa`) shows more tortuous paths; the right (`condition_2`, high `kappa`) shows straighter, more directed tracks. The mean resultant arrow is computed independently per panel.

## Advanced: Accessing Underlying Data

```{r trajset-advanced}
head(ts@data)
head(ts@meta$trial_limits)
```
