---
title: "radiatR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{radiatR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
if (requireNamespace("pkgload", quietly = TRUE)) {
  # Always load from source so the vignette reflects the current development state.
  pkgload::load_all("..", export_all = FALSE, helpers = FALSE, quiet = TRUE)
} else if (requireNamespace("radiatR", quietly = TRUE)) {
  library(radiatR)
} else {
  stop("Package 'radiatR' not installed and 'pkgload' not available.")
}
library(ggplot2)
```

## Overview

`radiatR` streamlines the journey from raw tracking files to publication-ready plots for experiments conducted in circular arenas. The package includes:

-   utilities for importing paired landmark/track text files produced by popular tracking suites;
-   helpers for extracting trial limits and computing circular summary statistics;
-   polished `ggplot2` layers for drawing concentric guides and annotated paths.

## Import Pipeline

The package ships five arc = 0° example tracks that demonstrate the full import workflow. Landmark files (`_point01.txt`) record two pixel-space reference points per trial (arena centre and stimulus edge); track files (`_point02.txt`) contain the full xy trajectory.

```{r load-demo}
track_dir     <- system.file("extdata", "tracks", package = "radiatR")
manifest_path <- system.file("extdata", "P_lividus_trials.csv", package = "radiatR")

file_tbl  <- import_tracks(track_dir)
manifest  <- import_info(manifest_path)
file_tbl  <- load_tracks(file_tbl, manifest, track_dir)

ts_demo <- get_all_object_pos(file_tbl = file_tbl, track_dir = track_dir)
ts_demo
```

`get_all_object_pos()` reads each landmark/track pair, normalises coordinates to a unit circle (arena radius = 1), and returns a `TrajSet`. Trial metadata (arena radius, stimulus position, frame limits) is in `ts_demo@meta$trial_limits`.

## Full *P. lividus* Dataset

The package also provides `plividus`, a pre-computed `TrajSet` containing all 401 trajectories from the experiment across six stimulus arc-angle conditions (0°, 15°, 30°, 45°, 60°, 150°). Loading it is instant.

```{r load-full}
data(plividus)
plividus
```

## Plotting Trajectories

`radiate()` draws trajectories on a unit circle with concentric reference rings. Colouring by the `arc` factor shows how paths cluster by condition.

```{r plot-all}
radiate(plividus,
        group_col  = "trial_id",
        colour_col = "arc",
        show_labels = FALSE,
        show_arrow  = FALSE)
```

## Heading Overlays

The **crossing method** — projecting the vector between two concentric-ring crossings to the unit circle — assigns one directional heading per trial. `derive_headings()` with `return_coords = TRUE` returns both the heading angle and the inner-ring crossing position.

```{r headings}
hd <- derive_headings(plividus, rule = "crossing",
                      circ0 = 0.2, circ1 = 0.4,
                      return_coords = TRUE)
# add arc grouping for faceting
hd$arc <- factor(
  as.integer(sub("G1D_(\\d+)_.*", "\\1", hd$id)),
  levels = c(0, 15, 30, 45, 60, 150),
  labels = c("0°", "15°", "30°", "45°", "60°", "150°")
)
names(hd)[names(hd) == "id"] <- "trial_id"
head(hd[, c("trial_id", "arc", "heading", "x_inner", "y_inner")])
```

Overlaying the heading endpoints (one hollow circle per trial) and the grand mean direction on the combined trajectory plot gives a compact summary of the full dataset:

```{r headings-overlay}
summ_all <- do.call(rbind, lapply(c(1), function(.) {
  h <- hd$heading
  data.frame(mean_dir    = atan2(mean(sin(h), na.rm = TRUE),
                                 mean(cos(h), na.rm = TRUE)),
             resultant_R = sqrt(mean(cos(h), na.rm = TRUE)^2 +
                                mean(sin(h), na.rm = TRUE)^2))
}))
seg_all <- circ_mean_segments(summ_all)

p_all <- radiate(plividus,
                 group_col   = "trial_id",
                 colour_col  = "arc",
                 show_labels = FALSE,
                 show_arrow  = FALSE) +
  add_heading_points(hd, colour_col = "arc", size = 1, alpha = 0.6)

gg_add_circ_mean(p_all, seg_all)
```

The grand mean arrow points at `r round(summ_all$mean_dir * 180 / pi, 1)`° with *R* = `r round(summ_all$resultant_R, 2)`, reflecting a moderate overall tendency across all conditions.

## Colour Options

Three strategies control how trajectories are coloured.

**Option 1 — single colour.** Pass a colour string directly to the heading-overlay helpers; omit `colour_col` from `radiate()` to draw all tracks in the default grey.

**Option 2 — cycling palette.** `colour_cycle` assigns each trajectory a colour index that cycles back to 1 after every *n* trajectories. When `panel_by` is set the cycle restarts independently within each panel so that trajectory 1 in every panel always gets colour 1. Pass an integer for automatic palette assignment, or a character vector to specify colours explicitly.

**Option 3 — variable mapping.** `colour_col` maps any data column to colour (e.g. the `arc` factor used in the combined-trajectory plot above).

## Per-Condition Panels

Faceting by arc angle shows each condition separately. Within each panel, `colour_cycle = 10` distinguishes individual trajectories — more informative than colouring by arc when every panel already corresponds to a single arc value. Per-condition mean arrows are computed from the crossing headings and routed to the correct panel automatically via the shared `arc` column.

```{r arc-panels, fig.width=8, fig.height=5}
arc_summ <- do.call(rbind, lapply(split(hd, hd$arc), function(sub) {
  h <- sub$heading
  data.frame(
    arc         = sub$arc[1],
    mean_dir    = atan2(mean(sin(h), na.rm = TRUE), mean(cos(h), na.rm = TRUE)),
    resultant_R = sqrt(mean(cos(h), na.rm = TRUE)^2 + mean(sin(h), na.rm = TRUE)^2)
  )
}))
segs <- circ_mean_segments(arc_summ)

radiate(plividus,
        group_col    = "trial_id",
        colour_cycle = 10,
        panel_by     = "arc",
        ncol         = 3,
        show_labels  = FALSE,
        show_arrow   = FALSE) +
  add_heading_vectors(hd, colour = "grey40") +
  add_heading_points(hd, colour = "grey40", size = 2) +
  geom_segment(
    data        = segs,
    aes(x = x, y = y, xend = xend, yend = yend),
    arrow       = grid::arrow(length = grid::unit(0.2, "cm")),
    colour      = "black",
    linewidth   = 1,
    inherit.aes = FALSE
  )
```

Each panel is labelled with its arc condition (strip above each panel), shows all trajectories coloured by cycling palette, the crossing heading vectors (dotted lines, grey), and a solid arrow whose direction and length represent the within-condition circular mean direction and resultant *R*. Use `strip_position = "inside"` to place the label inside the plot area, or `strip_labels = FALSE` to suppress it.

## Circular Density Overlays

Three functions handle directional density overlays; they are designed to compose cleanly with any density source:

| Function | Role |
|---|---|
| `compute_circular_density()` | Estimates density from raw headings; returns a plain data frame of `(theta, density)` pairs |
| `add_circular_density()` | Renders any `(theta, density)` data frame as a radial overlay — agnostic to how the density was produced |
| `add_heading_density()` | Convenience wrapper: calls the two above in sequence |

Three built-in methods are available in `compute_circular_density()`: `"vonmises"` (MLE via `circular::mle.vonmises()`), `"kernel"` (circular KDE), and `"histogram"` (bin counts). Because the computation and rendering steps are separate, the `density` column can be replaced with values from any other model — including a Bayesian posterior predictive density from `brms` — before calling `add_circular_density()`:

```{r density-concept, eval=FALSE}
dens_df <- compute_circular_density(hd, colour_col = "arc", method = "vonmises")
# replace with Bayesian posterior: dens_df$density <- my_brms_fitted_density(dens_df$theta)
add_circular_density(dens_df, colour_col = "arc", fill = "grey80", alpha = 0.35)
```

The convenience wrapper `add_heading_density()` skips the intermediate step when no substitution is needed. The combined panel plot below uses it to shade a per-condition von Mises density alongside the heading vectors:

```{r density-overlay, fig.width=8, fig.height=5}
radiate(plividus,
        group_col    = "trial_id",
        colour_cycle = 10,
        panel_by     = "arc",
        ncol         = 3,
        show_labels  = FALSE,
        show_arrow   = FALSE) +
  add_heading_density(hd, colour_col = "arc",
                      method = "vonmises", scale = 0.4,
                      fill = "grey80", alpha = 0.35) +
  add_heading_vectors(hd, colour = "grey40") +
  geom_segment(
    data        = segs,
    aes(x = x, y = y, xend = xend, yend = yend),
    arrow       = grid::arrow(length = grid::unit(0.2, "cm")),
    colour      = "black",
    linewidth   = 1,
    inherit.aes = FALSE
  )
```

The shaded region is the von Mises density fitted to the crossing headings within each arc condition. A narrower, taller peak indicates more concentrated directional responses. Switch to `method = "kernel"` for a non-parametric estimate, or `method = "histogram"` for a raw count display.

## Bootstrap Confidence Band

`compute_circular_density()` with `boot_reps > 0` runs a non-parametric bootstrap: heading samples are drawn with replacement, a von Mises MLE is fitted to each replicate, and the density is re-evaluated on the same angular grid. The `boot_alpha / 2` and `1 - boot_alpha / 2` quantiles across replicates become `density_lower` and `density_upper` columns, which `add_circular_density()` renders as a shaded band around the fitted curve.

```{r bootstrap-ci, fig.width=8, fig.height=5}
dens_boot <- compute_circular_density(hd, colour_col = "arc",
                                      method = "vonmises",
                                      boot_reps = 499L, boot_alpha = 0.05,
                                      n_theta = 300L)

radiate(plividus,
        group_col    = "trial_id",
        colour_cycle = 10,
        panel_by     = "arc",
        ncol         = 3,
        show_labels  = FALSE,
        show_arrow   = FALSE) +
  add_circular_density(dens_boot, colour_col = "arc",
                       scale = 0.4, fill = "grey80", alpha = 0.35,
                       ci_fill = "grey60", ci_alpha = 0.35) +
  geom_segment(
    data        = segs,
    aes(x = x, y = y, xend = xend, yend = yend),
    arrow       = grid::arrow(length = grid::unit(0.2, "cm")),
    colour      = "black",
    linewidth   = 1,
    inherit.aes = FALSE
  )
```

The darker band is the 95% bootstrap confidence interval for the fitted von Mises density. A wider band reflects greater parametric uncertainty, typically seen in the smaller or more diffuse conditions. The `density_lower` and `density_upper` columns in `dens_boot` can be replaced with interval values from a Bayesian model (e.g. the 2.5th and 97.5th percentiles of a `brms` posterior predictive distribution) before passing to `add_circular_density()`.

## Circular Summary Statistics

`circ_summary_headings()` provides the per-condition statistics behind the arrows above:

```{r stats}
arc_summ[, c("arc", "mean_dir", "resultant_R")]
```

For within-trial path consistency (tortuosity), `circ_summary()` operates on the step-by-step angle distribution:

```{r within-trial}
circ_summary(ts_demo)
```

High resultant lengths (close to 1) indicate very consistent step directions within a trial.

## Simulating Demo Data

`simulate_tracks()` generates synthetic trajectories without any files, useful for testing and demonstrations. The `kappa` parameter controls directedness:

```{r simulate}
sim_df <- simulate_tracks(
  conditions = data.frame(n_trials = c(5L, 5L), kappa = c(2, 8)),
  n_points = 150,
  seed = 42
)
ts_sim <- TrajSet(
  sim_df,
  id = "trial_id", time = "frame",
  angle = "rel_theta", x = "rel_x", y = "rel_y",
  angle_unit = "radians", normalize_xy = FALSE
)
radiate(ts_sim,
        group_col    = "trial_id",
        colour_cycle = 5,
        panel_by     = "condition",
        ncol         = 2,
        show_arrow   = TRUE)
```

The left panel (low `kappa`) shows tortuous paths; the right (high `kappa`) shows straighter, more directed tracks. The mean resultant arrow is computed independently per panel.
