---
title: "radiatR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{radiatR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(radiatR)
library(ggplot2)
```

## Overview

`radiatR` streamlines the journey from raw tracking files to publication-ready plots for experiments conducted in circular arenas. The package includes:

-   utilities for importing paired landmark/track text files produced by popular tracking suites;
-   helpers for extracting trial limits and computing circular summary statistics;
-   polished `ggplot2` layers for drawing concentric guides and annotated paths.

## Simulating Demo Data

Large tracking files are rarely suitable for inclusion in a package. To make example workflows reproducible, `simulate_tracks()` creates tidy demo data with controllable tortuosity.

```{r simulate}
demo_tracks <- simulate_tracks(n_trials = 4, n_points = 150, seed = 42)
head(demo_tracks)
```

The resulting tibble already contains relative coordinates (`rel_x`, `rel_y`) and angular summaries, so we can visualise it directly.

```{r radiate}
radiate(
  demo_tracks,
  x_col = "rel_x",
  y_col = "rel_y",
  group_col = "trial_id",
  colour_col = "tortuosity"
)
```

## File-Based Ingest Pipeline

When you have landmark/track text files on disk, the import helpers stitch them into tidy data structures. The example below creates a minimal synthetic data set to illustrate the workflow without requiring external assets.

```{r ingest}
tmp_dir <- tempfile("radiatr_ingest")
dir.create(tmp_dir)

# Write paired landmark/track files for a single video with two trials
landmarks <- data.frame(
  frame = c(1, 1, 120, 120),
  x = c(0, 40, 0, -35),
  y = c(0, 0, 0, 0)
)
tracks <- data.frame(
  frame = 1:200,
  x = c(seq(0, 25, length.out = 120), seq(25, -10, length.out = 80)),
  y = c(seq(0, 35, length.out = 120), seq(35, -5, length.out = 80))
)

write.table(
  landmarks,
  file = file.path(tmp_dir, "demo_video_point01.txt"),
  sep = "\t", col.names = FALSE, row.names = FALSE
)
write.table(
  tracks,
  file = file.path(tmp_dir, "demo_video_point02.txt"),
  sep = "\t", col.names = FALSE, row.names = FALSE
)

# Import metadata and enrich it with trial annotations
file_tbl <- import_tracks(tmp_dir)
manifest <- data.frame(
  file = file_tbl$basename,
  arc = c(0),
  type = "demo",
  obstacle = "none",
  id = "subject_01",
  stringsAsFactors = FALSE
)
file_tbl <- load_tracks(file_tbl, manifest, tmp_dir)

# Derive trial limits and normalised coordinates
landmarks_df <- read.delim(file.path(tmp_dir, file_tbl$landmark[1]),
                           header = FALSE)[, 1:3]
names(landmarks_df) <- c("frame", "x", "y")
tracks_df <- read.delim(file.path(tmp_dir, file_tbl$track[1]),
                        header = FALSE)[, 1:3]
names(tracks_df) <- c("frame", "x", "y")

trial_limits <- get_trial_limits(landmarks_df, tracks_df, file_tbl, vid_num = 1)
trial_objects <- get_tracked_object_pos(trial_limits, tracks_df)

str(trial_objects[[1]][[1]])
```

The list returned by `get_tracked_object_pos()` contains one tibble per trial as well as the augmented `trial_limits` data frame. These outputs feed directly into `radiate()` or any downstream statistical workflow.

```{r ingest-plot}
all_trial_tracks <- do.call(rbind, trial_objects[[1]])
radiate(
  all_trial_tracks,
  x_col = "rel_x",
  y_col = "rel_y",
  group_col = "trial_num",
  colour_col = "trial_num",
  label_col = "trial_num",
  show_arrow = TRUE
)
```

Remember to clean up when you are finished:

```{r cleanup, echo=FALSE}
unlink(tmp_dir, recursive = TRUE)
```