---
title: "Use Animal Tracks"
author: "John Kirwan"
date: "30/6/2021"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Get data from animal tracks

```{r message=FALSE, warning=FALSE}
library('ggplot2')
library('magrittr')
library('dplyr')
library('tidyr')
library('tibble')
library('circular')
library('ggforce')
```

Uses stimulus centre and point opposite stimulus centre as markers for the unit circle on arena. Gets complete track from each set of trials, makes a plot and determines headings.

```{r}
#folder <- 'D:/sci/2021/P_lividus_adult_vids/tracking/'
#folder <-  paste0(getwd(),'/tracks/')
folder <- "C:/Users/seano/Documents/2021/tracking/"
filez  <- list.files(normalizePath(folder),recursive = TRUE)
file_tbl <-tibble(basename=sub("*_point01.txt","",#subset str before exp
            list.files(normalizePath(folder),recursive = TRUE)[
            grep('*point01.txt$',list.files(normalizePath(folder),recursive=T))]))

file_tbl %<>% mutate(landmark=paste0(basename,"_point01.txt"),
                    track    =paste0(basename,"_point02.txt"))
tail(file_tbl)
```

Define the radius of the inner circle.

**NB:** As the tracks are uncalibrated, the size of objects in the central region are expanded relative to those at the periphery. This should become much more eccentric at the very edges and effect little change near the centre. The inner and outer radii are therefore somewhat smaller than the stated values. 

The radius criterion decides whether the first point with a radius larger than the specified inner and outer radii is used or whether the nearest point either side of the radius is used.

```{r}
circ0 <- .2#.25#.2 # proportion of inner radius covered by inner pt
circ1 <- .4 #.5#.4 # proportion of inner radius covered by outer pt
radius_criterion   <- "first_past" # "nearest"
stim_mid_criterion <- "mid_inflection" #"dark_trough" # #"dark_trough" #
```

Make functions to bring angle values back into the fold.

```{r}
rad2clock   <- function(theta){theta<-(pi/2)-theta
                if(theta < 0){theta <- theta + 2*pi
                }else{ if(theta > 2*pi){theta <- theta - 2*pi} }
                return(theta)}

rad_unclock <- function(theta){theta<-(pi/2)-theta
                if(theta < -pi){theta <- theta + 2*pi
                }else{ if(theta > pi){theta <- theta - 2*pi} }
                return(theta)}

rad_shepherd <- function(theta){  if(theta < -pi){theta <- theta + 2*pi
} else{ if(theta > pi){theta <- theta - 2*pi} }
  return(theta)}
```

#### Bring in data frame with video names and trial order

Obst means obstructed - there was initially an obstacle in the way.

```{r}
df <- readr::read_csv('./../data/Pl_adult_trials_list.csv')
df$obstacle[df$obstacle=="none"]   <-"unobst"
df$obstacle[df$obstacle=="marbles"]<-"obst"
df %<>% tidyr::unite("cond",c(type,arc,obstacle),remove =F,sep="_")
head(df)
if(any(!file_tbl$basename %in% df$file)){warning("Missing file")}
if(any(!df$file %in% file_tbl$basename)){warning("Missing file")}
```

Add part above to put trial, cond, type, arc, obstacle into file_tbl and therefore to trackz_tbl and trialz.

```{r}
# which track rows shares vid name with df
file_tbl$arc  <- rep(0,dim(file_tbl)[1])
file_tbl$type <- as.character(rep("",dim(file_tbl)[1]))
file_tbl$obstacle <- as.character(rep("",dim(file_tbl)[1]))
file_tbl$id       <- as.character(rep("",dim(file_tbl)[1]))

i=1
while(i <= dim(df)[1]){
 file_tbl$arc [which(file_tbl$basename==df$file[i])] <- df$arc[i]
 file_tbl$type[which(file_tbl$basename==df$file[i])] <- df$type[i]
 file_tbl$obstacle[which(file_tbl$basename==df$file[i])] <- df$obstacle[i]
 file_tbl$id[which(file_tbl$basename==df$file[i])] <- df$id[i]
 i=i+1      }

file_tbl  %<>% tidyr::unite("cond",c(type,arc,obstacle),remove=FALSE, sep="_")
```

Are there any (predicted) track names not represented in the file list?
 i.e. landmark files without a matching track file.

```{r}
if(any(!file_tbl$track %in% list.files(normalizePath(folder),recursive = TRUE))){
  stop(print(paste0("The following track file is missing:",
    file_tbl$track[which(!file_tbl$track %in% list.files(
      normalizePath(folder),recursive = TRUE))])))                            }

# are any track files in the file list without a match in the landmark files?
if(any(!list.files(normalizePath(folder),recursive = TRUE)[grep('*point02.txt$',
    list.files(normalizePath(folder),recursive = TRUE))] %in% file_tbl$track)){
  stop(print(paste0("The following track file lacks a counterpart:",
    list.files(normalizePath(folder),recursive = TRUE)[grep('*point02.txt$',
    list.files(normalizePath(folder),recursive = TRUE))])))                   }
```

Try with first file to get the track

```{r}
animal_track = read.delim(normalizePath(paste0(folder,'/',file_tbl$track[1])),
                          sep="\t", header = FALSE)[,1:3]
names(animal_track) <- c("frame","x","y")
head(animal_track)
```

Try with first file to get the landmarks

```{r}
landmarks = read.delim(normalizePath(paste0(folder,'/',file_tbl$landmark[1])),
                          sep="\t", header = FALSE)[,1:3]
names(landmarks) <- c("frame","x","y")
head(landmarks)

if(length(landmarks$x) %% 2 == 1){ warning("Odd number of landmarks!") }
num_trials = length(landmarks$x)/2 
```

## The files 

With the assignment pipe *%<>%*, the variable is assigned to itself at the end. The radius of the circle based on the stimulus distance is *r_px*. atan2 outputs angles in radians according to the unit cicle from -pi to pi.

The y values of the landmark and tracks have been reversed to run upwards rather than downwards.

*stim_x_0* and *stim_y_0* are cartesian coords which have there location translated by subtracting the origin and the length scaled to the distance from the origin (and points transformed accordingly).

```{r}
get_trial_limits <- function(landmarks,animal_track,file_tbl,vid_num){ #,

 num_trials = dim(landmarks)[1]/2 # count number of trials
 # pick out start frame and centre and stimulus coords
 tl <- as_tibble(array(0,dim=c(num_trials,18),
    dimnames = list(NULL, c("first_f","orig_x","orig_y","stim_x","stim_y",
    "last_f","stim_x_0","stim_y_0","stim_theta","r_px","quadrant","video",
    "order","vid_ord","x0","y0","x1","y1"))),  .name_repair="unique")
  tl$video <- as.character(tl$video)
  landmarks$y <- - landmarks$y # invert y axis to read from bottom to top
  
  i = 1 
 while (i <= num_trials){ # for each trial
  tl$first_f[i]<- as.numeric(landmarks[(i*2)-1,1])# trial i start frame
  tl$orig_x[i] <- as.numeric(landmarks[(i*2)-1,2])# trial i centre pt x,y
  tl$orig_y[i] <- as.numeric(landmarks[(i*2)-1,3])# trial i centre pt x,y
  tl$stim_x[i] <- as.numeric(landmarks[i*2,2])# trial i stim centre x,y
  tl$stim_y[i] <- as.numeric(landmarks[i*2,3])# trial i stim centre x,y
       if (i < num_trials){
        tl$last_f[i] <- as.numeric(landmarks[((i+1)*2)-1,1]-1)# trial i last frame
     } else { tl$last_f[i]  <- max(animal_track[,1]) } # max along dimension 1 
  i = i + 1
  }
  
 tl$video    <- file_tbl[[vid_num,1]]
 tl$arc      <- file_tbl$arc[file_tbl$basename==tl$video]
 tl$type     <- file_tbl$type[file_tbl$basename==tl$video]
 tl$obstacle <- file_tbl$obstacle[file_tbl$basename==tl$video]
 tl$id       <- file_tbl$id[file_tbl$basename==tl$video]
 
 tl$stim_x_0 <- tl$stim_x - tl$orig_x # stimulus cart coords - orig
 tl$stim_y_0 <- tl$stim_y - tl$orig_y
 tl$stim_theta <- atan2(tl$stim_y_0,tl$stim_x_0) 
 tl$r_px     <- sqrt((tl$stim_x_0)^2+(tl$stim_y_0)^2) # pixel radius
 # if the stimulus centre is set to the midpoint, shift all stimulus positions accordingly
 if(stim_mid_criterion=="mid_inflection" & tl$type[1]=="Herm"){    # shift .5 T clockwise
   # subtracting angles only works in +ve space
   stim_theta    <- mapply(rad2clock, tl$stim_theta) # change to clock
   stim_theta    <- stim_theta - tl$arc*(pi/360) # add T/2 to shift clockwise 
   stim_theta    <- mapply(rad_shepherd_clock,stim_theta) #herd into 0 to 2*pi
   tl$stim_theta <- mapply(rad_unclock,stim_theta) # back to unit circle
 }
 ##### MAY NEED TO REMAKE THESE BELOW!
 #tl$stim_x_0 <- tl$stim_x_0 / tl$r_px # standardize by the radius 
 #tl$stim_y_0 <- tl$stim_y_0 / tl$r_px # standardize by the radius 
 tl$stim_x_0  <- cos(tl$stim_theta)
 tl$stim_y_0  <- sin(tl$stim_theta)
 
 i=1
 while (i <= num_trials){ 
  if (tl$stim_theta[i] > .25*pi & tl$stim_theta[i] <= .75*pi){
      tl$quadrant[i] = "top" #if stimulus in first quadrant
  }else if (tl$stim_theta[i]>= -.25*pi & tl$stim_theta[i] < .25*pi){
      tl$quadrant[i] = "left"
  }else if (tl$stim_theta[i] >= -.75*pi & tl$stim_theta[i] < -.25*pi ){
      tl$quadrant[i] = "bottom"
  }else {
      tl$quadrant[i] <- "right" } 
  i = i + 1
 }
 return(tl)
}
```

Find each trial and get data, (i) frame no (ii - v) coords for centre and stimulus (vi) last frame, (vii - viii) normalized stimulus centre (ix) theta in rads (anticlock, zero=0) (x) rho in pixels 

Get limits from first video 

```{r}
trial_limits <- get_trial_limits(landmarks,animal_track,file_tbl,1) #
```


#### Get animal tracks in polar terms relative to the stimulus

Divide the animal tracks up by trial using the first landmark. Initialize 
transformed tracks for later.

```{r Get animal pos fx}
get_animal_pos <- function(trial_limits,animal_track){
  num_trials = dim(trial_limits)[1] # count number of trials
  
  animal_track$y <- -animal_track$y # invert y so that runs from bottom up
  # make nested tibble of the sets
  trackz <- purrr::map(vector(length=num_trials), tibble)

 i <- 1 # initialize i  
 while (i <= dim(trial_limits)[1]){    # for each trial in a video 

   animal_track$trial_num[ animal_track$frame >= trial_limits$first_f[i] & 
     animal_track$frame <= trial_limits$last_f[i]] <- i # get the trial no for each obs
   XY <- as_tibble(animal_track[animal_track$frame >= trial_limits$first_f[i] & 
        animal_track$frame <= trial_limits$last_f[i],]) # make a tibble of that trial
   XY %<>% add_column(trans_x=0,trans_y=0,abs_theta=0,trans_rho=0,video=NA,
                      order=NA,vid_ord=NA,rel_theta=0)
   XY[,5:6] <- sweep(XY[,2:3],2,as.numeric(trial_limits[i,2:3]),"-") # subtract origin
   XY[,5:6] <- XY[,5:6] / trial_limits$r_px[i]
   
   # get theta from Cartesian coordinates
   abs_theta    <- mapply(atan2,XY$trans_y,XY$trans_x)
   # reorient abs theta to be clockwise and with a zero at the top
   XY$abs_theta <- mapply(rad2clock,abs_theta)

   # looks like this line below should apply to rel_theta
   XY$trans_rho <- unlist(sqrt(XY[,5]^2 + XY[,6]^2)) #/as.numeric(trial_limits$r_px[i]) #normalised rho
   XY$video <- trial_limits$video[i]
   XY$order <- as.character(i) # order of trials in video

   XY %<>% tidyr::unite("vid_ord", video:order,remove = FALSE, sep="_")
   stim_theta <- rad2clock(trial_limits$stim_theta[i])
   # reorient stim_theta and get relative position
   XY$rel_theta <- XY$abs_theta - stim_theta
   XY$rel_theta <- mapply(rad_shepherd_clock,XY$rel_theta) # herd back into clock
   XY$rel_theta <- mapply(rad_unclock,XY$rel_theta)# get back between -pi and pi
   # get absolute theta values back too
   XY$abs_theta <- mapply(rad_unclock,XY$abs_theta) # return to unit circle
   XY$rel_x     <- XY$trans_rho*cos(XY$rel_theta)
   XY$rel_y     <- XY$trans_rho*sin(XY$rel_theta)
   trackz[[i]]  <- XY
   # assign tracks less than 40% of arena to be empty
   if(min(XY$trans_rho>.4)){trackz[[i]] <- "Track starts too far from centre"
    warning(paste('Track starts too far from centre:'),XY$video[1])}
   if(any(XY$trans_rho> 1)){
    warning(paste('Track exceeds arena width: '),XY$video[1])}
   
   trial_limits$order[i]   <- as.character(i) 
   trial_limits$vid_ord[i] <- paste0(trial_limits[i,12],"_",i)
   

  if(radius_criterion=="first_past"){ # if taking 1st point outside radii
  if(!any(XY$trans_rho > circ0)){warning('NO PTS BEYOND INNER CIRC!!')} 
  trial_limits$x0[i] <- XY$trans_x[XY$trans_rho >= circ0][1] #get 1st pts 
  trial_limits$y0[i] <- XY$trans_y[XY$trans_rho >= circ0][1] #beyond inner circ
  # if pts both sides of outer circ
  if(any(XY$trans_rho <= circ1) & any(XY$trans_rho > circ1)){ 
    trial_limits$x1[i] <- XY$trans_x[XY$trans_rho >= circ1][1]
    trial_limits$y1[i] <- XY$trans_y[XY$trans_rho >= circ1][1]
  }else if(!any(XY$trans_rho > circ1)){warning("NO POINT BEYOND OUTER RADIUS!!!")
    trial_limits$x1[i] <- XY$trans_x[which.max(XY$trans_rho)] # get most distant point
    trial_limits$y1[i] <- XY$trans_y[which.max(XY$trans_rho)] # get most distant point
  }else{warning("NO POINT WITHIN OUTER RADIUS!!!")
    trial_limits$x1[i] <- XY$trans_x[i+1] # take the 2nd points outside
    trial_limits$y1[i] <- XY$trans_y[i+1] # as x0,y0 will be 1st
  }
  }else{
    trial_limits$x0[i] <- XY$trans_x[which.min(abs(XY$trans_rho - circ0))]
    trial_limits$y0[i] <- XY$trans_y[which.min(abs(XY$trans_rho - circ0))]
    trial_limits$x1[i] <- XY$trans_x[which.min(abs(XY$trans_rho - circ1))] #.5
    trial_limits$y1[i] <- XY$trans_y[which.min(abs(XY$trans_rho - circ1))]
  }

   i = i + 1
 }
 return(list(trackz,trial_limits)) # norm to radius
}
```

Get the all the tracks from a single file. Print one set.

```{r}
trial_track_list   <- get_animal_pos(trial_limits,animal_track)
trackz             <- trial_track_list[[1]]
tail(trackz[[1]])
```

```{r}
trial_limits      <- trial_track_list[[2]]
tail(trial_limits)
```

Now apply this to all files.Then transform the coordinates from all tibbles in the big list.

```{r}
get_all_animal_pos <- function(landmarks,animal_track,file_tbl){
  i <- 1 # for all files in file_tbl
  while(i <= dim(file_tbl)[1]){
    animal_track = read.delim(normalizePath(paste0(folder,'/',file_tbl$track[i])),
                            sep="\t", header = FALSE)[,1:3]
    names(animal_track) <- c("frame","x","y")
    landmarks <- read.delim(normalizePath(paste0(folder,'/',
                      file_tbl$landmark[i])),sep="\t", header = FALSE)[,1:3]
    names(landmarks) <- c("frame","x","y")
    if(length(landmarks$x) %% 2 == 1){
      warning(paste("Odd number of landmarks:",file_tbl$basename[i])) }
    trial_limits <- get_trial_limits(landmarks,animal_track,file_tbl,i) #
    trial_track_list <- get_animal_pos(trial_limits,animal_track) # run get_animal_pos
     if(i == 1){  
       all_trackz       <- trial_track_list[[1]] # populate with first list
       trialz <- trial_track_list[[2]] # first instance from above
     } else {
       all_trackz       <- c(all_trackz,trial_track_list[[1]])
       trialz <- rbind(trialz,trial_track_list[[2]])
    }
    i <- i + 1
  }
 trackz_n_limits <- list(all_trackz,trialz) 
 return(trackz_n_limits)
}
```


```{r}
trackz_n_limits  <- get_all_animal_pos(landmarks,animal_track,file_tbl)
all_trackz       <- trackz_n_limits[[1]]
tail(all_trackz[[1]])
```

```{r}
trialz <- trackz_n_limits[[2]]
tail(trialz)
```

#### Find cases where the inner and outer radii don't differ

This has stopped working to remove those trials - that's OK as identifying them is more important.

```{r}
idx <- which(trialz$x0==trialz$x1 & trialz$y0==trialz$y1)
if(length(idx) > 0){warning(paste("Trials",idx,"have errors!"))}
#all_trackz[idx] <- NULL
##### LEAVE OUT FOR MOMENT!
#trialz <- trialz[-c(idx),] 
## something wrong with this line!
print(paste(length(all_trackz),"tracks post culling"))
print(paste(dim(trialz)[1],"trials in limits list post culling"))
```

Make tibble for tracks rather than list

```{r}
trackz_tbl <- purrr::map_dfr(all_trackz, ~ .x)
```


#### Line-Circle Intercept.

Works when vector intersects circle.

```{r}
linecirc <- function(x0,y0,x1,y1){
  h = 0;k=0;radius=1 # circle origin x,y
  a = (x1 - x0)^2 + (y1 - y0)^2 # vector length
  b = (2*(x1 - x0)*(x0 - h)) + (2*(y1 - y0)*(y0 - k)) # 
  c = (x0 - h)^2 + (y0 - k)^2 - radius^2
  D <- (b^2) - (4*a*c) # discriminant - must be positive
  if(a <= 0){warning("Vector length is not positive!")}
  if(D <= 0){warning(paste0(
    "Discriminant is not positive with x0=",x0," y0=",y0," x1=",x1," y1=",y1))}
  if(c >= 0){warning("Starting pt is outside radius!")}
  
  # the roots for t may be found in the usual quadratic form:
  t <- c(0,0)
  t[1] = (-b + sqrt(D)) / (2*a)
  t[2] = (-b - sqrt(D)) / (2*a) 
  # 1st and 2nd rows of ints correspond to 1st and 2nd root respectively
  ints <- rbind(unlist(c( t[1]*(x1 - x0) + x0, t[1]*(y1 - y0) + y0)),
                unlist(c( t[2]*(x1 - x0) + x0, t[2]*(y1 - y0) + y0)))
            
  d <- c(sqrt(ints[1,1]^2 + ints[1,2]^2),sqrt(ints[2,1]^2 + ints[2,2]^2))
  if(all.equal(d[1],d[2])==FALSE){warning("Intersect is off unit circle!")}
  # 1st and 2nd rows of dist2track correspond to 1st and 2nd root respectively
  dist2track    <- c(0,0) 
  dist2track[1] <- sqrt( (ints[1,1]-x1)^2 + (ints[1,2]-y1)^2 )
  dist2track[2] <- sqrt( (ints[2,1]-x1)^2 + (ints[2,2]-y1)^2 )
  xy <- ints[which.min(dist2track),]
  
  # commented because NAs crash this part
  #if(t[1] < 1 | t[2] > 0){ #which.min(dist2track)
  #  warning('Roots imply vector is weird') }
  return(tibble::tibble(x_int = xy[1],y_int = xy[2])) # return closest to outer pt
}
```

Get the headings from intersection with the unit circle.

Use atan2 to get the headings in radians. Then, having changed the class to circular, set them so that 0 refers to the stimulus midpoint and the rotation is clockwise. 

The value *degs* differs from *radians* in beginning at the top of the circle and continuing clockwise from 0 to 360. 

```{r}
intz <- purrr::pmap_dfr(list(trialz$x0,trialz$y0,trialz$x1,trialz$y1), linecirc)

trialz$x_int    <- intz$x_int
trialz$y_int    <- intz$y_int

trialz$abs_rads <- mapply(atan2,trialz$y_int,trialz$x_int)
abs_rads_clock  <- mapply(rad2clock,trialz$abs_rads) # convert to be 0 to 2*pi, clockwise   
clockstim       <- mapply(rad2clock,trialz$stim_theta) # stim theta to clock
rel_radians     <- mapply(rad_shepherd_clock,abs_rads_clock - clockstim) #herd in clock
trialz$radians  <- mapply(rad_unclock,rel_radians) # back to unit circle

# degree clockwise variable for the trials and tracks
trialz$degs     <- (180/pi)*mapply(rad2clock,trialz$radians)
trackz_tbl$degs <- (180/pi)*mapply(rad2clock,trackz_tbl$rel_theta)
```

#### Make variables for *successful* orientation to target sector 

Then, plug these into the tracks table. 

```{r}
trialz$succ4 <- rep(FALSE,dim(trialz)[1])
trialz$succ5 <- rep(FALSE,dim(trialz)[1])
trialz$succ6 <- rep(FALSE,dim(trialz)[1])
trialz$succ4[trialz$degs<=45|trialz$degs>315]<-TRUE
trialz$succ5[trialz$degs<=36|trialz$degs>324]<-TRUE
trialz$succ6[trialz$degs<=30|trialz$degs>330]<-TRUE

trackz_tbl$succ4 <- rep(FALSE,dim(trackz_tbl)[1])
trackz_tbl$succ5 <- rep(FALSE,dim(trackz_tbl)[1])
trackz_tbl$succ6 <- rep(FALSE,dim(trackz_tbl)[1])

i = 1
while(i <= dim(trialz)[1]){
  trackz_tbl$succ4[trackz_tbl$vid_ord==trialz$vid_ord[i]] <- trialz$succ4[i]
  trackz_tbl$succ5[trackz_tbl$vid_ord==trialz$vid_ord[i]] <- trialz$succ5[i]
  trackz_tbl$succ6[trackz_tbl$vid_ord==trialz$vid_ord[i]] <- trialz$succ6[i]
  i=i+1
}
```

Temporary fix - remove observations with rho value greater than 1. Remove cases where there were more than 4 trials in the same video.

```{r}
trackz_tbl <- trackz_tbl[trackz_tbl$trans_rho <= 1,]
trackz_tbl <- trackz_tbl[!trackz_tbl$order > 4,]
trialz <- trialz[!trialz$order > 4,]
```


Check which trails are also in df

```{r}
# which track rows shares vid name with df
trackz_tbl$arc  <- rep(0,dim(trackz_tbl)[1])
trackz_tbl$type <- as.character(rep("",dim(trackz_tbl)[1]))
trackz_tbl$obstacle <- as.character(rep("",dim(trackz_tbl)[1]))
trackz_tbl$id <- as.character(rep("",dim(trackz_tbl)[1]))

trialz$arc  <- rep(0,dim(trialz)[1])
trialz$type <- as.character(rep("",dim(trialz)[1]))
trialz$obstacle <- as.character(rep("",dim(trialz)[1]))

i=1
while(i <= dim(df)[1]){
 trackz_tbl$arc [which(trackz_tbl$video==df$file[i])] <- df$arc[i]
 trackz_tbl$type[which(trackz_tbl$video==df$file[i])] <- df$type[i]
 trackz_tbl$obstacle[which(trackz_tbl$video==df$file[i])] <- df$obstacle[i]
 trackz_tbl$id[which(trackz_tbl$video==df$file[i])] <- df$id[i]
 trialz$arc [which(trialz$video==df$file[i])]     <- df$arc[i]
 trialz$type[which(trialz$video==df$file[i])]     <- df$type[i]
 trialz$obstacle[which(trialz$video==df$file[i])] <- df$obstacle[i]
 i=i+1
}

trackz_tbl %<>% tidyr::unite("cond",c(type,arc,obstacle),remove=FALSE, sep=" ")
trialz     %<>% tidyr::unite("cond",c(type,arc,obstacle),remove=FALSE, sep=" ")
```

#### Make variables for inner radius relative to the stimulus

Make something close to stimulus relative versions of the inner circle intersects. Make angles +ve real and subtract the stimulus to get the relative angular position of the inner circle intersect. Then, turn into Cartesian coordinates.

```{r}
clockINtheta    <- mapply(rad2clock,atan2(trialz$y0,trialz$x0))
clockstim       <- mapply(rad2clock,trialz$stim_theta)
rel_circ0_theta <- mapply(rad_shepherd_clock,clockINtheta-clockstim)
rel_circ0_theta <- mapply(rad_unclock,rel_circ0_theta) 
trialz$rho0     <- sqrt(trialz$x0^2 + trialz$y0^2)
trialz$rel_x0   <- trialz$rho0*cos(rel_circ0_theta)
trialz$rel_y0   <- trialz$rho0*sin(rel_circ0_theta)
```


Make the arc width a factor. This doesn't seem to influence the plotting order and may be redundant.

```{r}
trackz_tbl <- within(trackz_tbl, cond <- factor(trackz_tbl$cond,levels=c(
  'Herm 0 obst','Herm 15 obst','Herm 30 obst','Herm 45 obst','Herm 60 obst',
  'Herm 150 obst','Herm 150 unobst')))
```

Make degree class circular

```{r}
trialz$degs <- as.circular(trialz$degs,type="angles",
    units="degrees",zero=pi/2,rotation="clock",template= 'none',modulo='asis')
```


#### Plotting

*Points are plotting correctly although the zero and rotation are according to the unit circle because the points themselves are oriented that way, e.g. an upward pointing track has a heading of c.1.6.*

```{r}
line_df <- data.frame(x= c(0,.66,.95,.66,0,-.66,-.95,-.66),y = c(.95,.66,0,-.66,-.95,-.66,0,.66),
 xend = c(0,.74,1.05,.74,0,-.74,-1.05,-.74),yend = c(1.05,.74,0,-.74,-1.05,-.74,0,.74) )
```


For some reason it appears to be using the conventional unit circle zero and rotation (points concentrated at x=1,y=0).

Later, use scale_colour_brewer(palette = "Paired")

#### Absolute position of tracks with regard to the arena (not the stimulus)

```{r abs headings fig, fig.height = 9.5, fig.width = 7}
how_many <- 150
# geom_path is already facing up, without (pi/2)- , geom_point() is not
trackz_tbl |> filter(obstacle=="obst") |> 
  ggplot() + coord_fixed() +
  geom_path(aes(x=trans_x,y=trans_y,group=id,color=id),
             alpha=.5, size=.75) + 
  geom_point(aes(x=trans_x, y=trans_y,group=id,color=id),
             alpha=.05, size=.5) + 
  facet_wrap(vars(arc),ncol=2) + # obstacle
  annotate("path", color="orange",
  x=0+1*cos(seq(0,2*pi,length.out=100)),y=0+1*sin(seq(0,2*pi,length.out=100))) +
  annotate("path", color="blue",linetype="dashed",
  x=0+circ1*cos(seq(0,2*pi,length.out=100)),
  y=0+circ1*sin(seq(0,2*pi,length.out=100))) +
  annotate("path", color="black",linetype="dashed",
  x=0+circ0*cos(seq(0,2*pi,length.out=100)),
  y=0+circ0*sin(seq(0,2*pi,length.out=100))) +
  geom_point(data=trialz[
    trialz$obstacle=="obst" & as.numeric(trialz$id) <= how_many,],   
    aes(x=x_int,y=y_int,group=cond,color=id),
             alpha=.1,size=6,shape=1) +
  geom_text(data=trialz[
    trialz$obstacle=="obst" & as.numeric(trialz$id) <= how_many,],
   aes(x=x_int,y=y_int,color=id,label=id),alpha=1,size=3) +
  geom_segment(data=trialz[
    trialz$obstacle=="obst" & as.numeric(trialz$id) <= how_many,],
   aes(x=x0,y=y0,xend=x_int,yend=y_int,alpha=0.5,colour=id)) +
  geom_segment(data=line_df,mapping=aes(x=x,y=y,xend=xend,yend=yend),inherit.aes=F)+
   xlab('') + ylab('') + theme_classic() + 
  labs(title='')+#Plot tracks with obstacle relative to stimulus') +
  theme(axis.line=element_blank(),axis.ticks=element_blank(),
        axis.text=element_blank()) + 
    theme(legend.position = "none") +
  theme(panel.spacing = unit(.5, "cm")) +
  theme(plot.title=element_text(size=14,hjust=0.5),#vjust=-62), 
        strip.background=element_blank(),strip.text=element_blank()) +
  theme(panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent')
  )
```

### Obstructed all absolute headings powerpoint

```{r rel headings all Obst abs sheet fig, fig.height = 5.625, fig.width = 10}
library('ggnewscale')
#trackz_tbl$arc <- factor(trackz_tbl$arc)
df$id <- factor(df$id)
obstruct <- "obst"#c("unobst","obst") #"obst" #
arc_angs <- c(0:150)

linetypez <- c("FALSE"="dashed","TRUE"="solid")
trackz_tbl |>  filter(obstacle %in% obstruct & arc %in% arc_angs) |> 
ggplot() + coord_fixed() + 
 #facet_wrap(vars(arc,obstacle),ncol=3,strip.position="top")+
 annotate("path",colour="black",x=cos(seq(0,2*pi,0.06)),y=sin(seq(0,2*pi,0.06)))+
 new_scale("color")+  
 geom_path(aes(x=trans_x, y=trans_y,color=factor(arc),group=vid_ord))+ 
 theme(legend.position = "none") +
 annotate("path", color="blue",linetype="dashed",
  x=circ1*cos(seq(0,2*pi,0.05)),
  y=circ1*sin(seq(0,2*pi,0.05))) +
 annotate("path", color="black",linetype="dashed",
  x=circ0*cos(seq(0,2*pi,0.05)),y=circ0*sin(seq(0,2*pi,0.05))) +
 #annotate("text",x=c(1.1,0,-1.1,0),y=c(0,1.1,0,-1.1),label=c("90","0","270","180"))+
 #annotate("text", x=c(1.1,0),y=c(0,1.1),label=c("90","0")) +  
 annotate("text", x= .8, y= .8, label=paste0('NW',""))  +  
 annotate("text", x= 0, y= 1.1, label=paste0('W',""))  +  
 annotate("text", x= .8, y=-.8, label=paste0('NE',"")) +  
 annotate("text", x= 0, y= -1.1, label=paste0('E',""))  +  
 annotate("text", x=-.8, y=-.8, label=paste0('SE',"")) +  
 annotate("text", x= -1.1, y= 0, label=paste0('S',""))  +  
 annotate("text", x=-.8, y= .8, label=paste0('SW',"")) +
 annotate("text", x=1.1, y= 0, label=paste0('N',"")) +
 geom_point(data=trialz[ # Stick on final relative headings
  trialz$obstacle==obstruct & trialz$arc %in% arc_angs & as.numeric(trialz$id) <= how_many,],
  aes(x=cos(abs_rads),y=sin(abs_rads),color=factor(arc)),size=2,shape=1)+ scale_shape_identity()+ 
 geom_segment(data=trialz[ # Subliminal R bar arrow 
   trialz$obstacle %in% obstruct & trialz$arc %in% c(0:15),],
   aes(x=0,y=0,xend=rho.circular(abs_rads)*cos(mean.circular(abs_rads)),
   yend=rho.circular(abs_rads)*sin(mean.circular(abs_rads))),size=1.3,
   colour="grey",arrow=arrow(length=unit(.2,"cm"))) +
 geom_segment(data=trialz[ # Superliminal R bar arrow 
   trialz$obstacle %in% obstruct & trialz$arc %in% c(40:60),],
   aes(x=0,y=0,xend=rho.circular(abs_rads)*cos(mean.circular(abs_rads)),
   yend=rho.circular(abs_rads)*sin(mean.circular(abs_rads))),size=1.3,
   colour="black",arrow=arrow(length=unit(.2,"cm"))) +  
  
 geom_segment(data=line_df,aes(x=x,y=y,xend=xend,yend=yend),inherit.aes=F) +
  xlab('') + ylab('') + theme_classic() + 
  labs(title='',color = "Width")+#Plot tracks with obstacle relative to stimulus') +
  theme(axis.line=element_blank(),axis.ticks=element_blank(),
        axis.text=element_blank()) + 
  theme(panel.spacing = unit(.5, "cm")) +
  theme(plot.title=element_text(size=14,hjust=0.5),#vjust=-62), 
        strip.background=element_blank(),strip.text=element_blank()) +
  theme(panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent',color=NA), #transparent legend bg
    legend.box.background = element_rect(fill='transparent',color=NA) #transparent legend panel
  )
ggsave('Plividus_prelim_abs_head_150d_unobst.svg',bg='transparent')
```

The grey arrow is represents the circular mean and mean result length of the two lowest levels (0d and 15d), whereas the black arrow represents the 45d and 60d stimuli. If there were much more directedness for the subliminal set, this might indicate that in the absence of a clear visual stimulus, they are drawn to other visual features in the environment. 


#### Are animals attracted to the top-right (the opening)?

From the plots, there may be attraction in some instances - check with null hypothesis tests.

```{r}
#trialz$succ <- rep(0,dim(trialz)[1])
trialz |> group_by(type,arc,obstacle) |>
  summarize( 
    n    = n(),
    ids  = n_distinct(id),
    mu_r = as.numeric((mean.circular(abs_rads))),
    R    = round(rho.circular(abs_rads),2),
    Ray.p= round(rayleigh.test(abs_rads)[[2]],2),
    V.p  = round(rayleigh.test(abs_rads,mu=circular(pi/4))[[2]],2), # to top-right
    #quart= round(sum(succ4)/length(succ4),2),
    #Binp4= round(binom.test(sum(succ4),length(succ4),p=1/4)[[3]],2),
    #quint= round(sum(succ5)/length(succ5),2),
    #Binp5= round(binom.test(sum(succ5),length(succ5),p=1/5)[[3]],2),
    #sext = round(sum(succ6)/length(succ6),2),
    #Binp6= round(binom.test(sum(succ6),length(succ6),p=1/6)[[3]],2),
    sd   = round(sd.circular(abs_rads),2)
             ) -> abs_summ
abs_summ
```

Yes, according to the V test animals are likely attracted to the opening with the negative grey control, which is perhaps, unsurprising as there was no competing stimulus. They were also attracted to the 60d Hermitian stimulus. (Could something have differed in those trials - could the stimulus have had reduced contrast?)

#### Relative position of tracks with regard to the stimulus

The trial limits radians and the tracks now align. 

```{r}
RhoCircularRad <- function(x) {
   n <- length(x)   
   sinr <- sum(sin(x))
   cosr <- sum(cos(x))
   result <- sqrt(sinr^2 + cosr^2)/n
   return(result)
}

makerad <- function(x){x <- as.circular(x,type="angles",units="radians", 
  template="none",modulo="asis",zero=0,rotation="clock")
  }
```


### Relative tracks - summary table of conditions

Add variable for successfully falling into target sextant.

```{r Summarize relative headings}
trialz |> group_by(type,arc,obstacle) |>
  summarize( 
    n    = n(),
    ids  = n_distinct(id),
    mu_d = as.numeric(round(mean.circular(degs))), # differs from radians
    R    = round(rho.circular(degs),3),            # same deg or radians
    Ray.p= round(rayleigh.test(degs)[[2]],3),
    V.p  = round(rayleigh.test(degs,mu=circular(pi/2))[[2]],3), # same deg or rad
    quart= round(sum(succ4)/length(succ4),2),
    Binp4= round(binom.test(sum(succ4),length(succ4),p=1/4)[[3]],2),
    quint= round(sum(succ5)/length(succ5),2),
    Binp5= round(binom.test(sum(succ5),length(succ5),p=1/5)[[3]],2),
    sext = round(sum(succ6)/length(succ6),2),
    Binp6= round(binom.test(sum(succ6),length(succ6),p=1/6)[[3]],2),
    sd   = round(sd.circular(degs),2),
    mu_r = mean.circular(radians), # differs from degrees
             ) -> headings_summ
#headings_summ$arc <- factor(headings_summ$arc)
headings_summ
```

```{r}
deg_labs <- rlang::expr(
annotate("text", x=0,y=1.1,label="0") +  
annotate("text", x=1.1,y=0,label="90") +  
annotate("text", x=0,y=-1.1,label="180") +  
annotate("text", x=-1.1,y=0,label="270") )
```


Uncalibrated tracks from the P. lividus adult trials. Each figure panel, represents the experiment arena looking downwards from above. The outer orange circle represents the wall of the arena, with is coloured grey, except for a visual stimulus, graded across the horizon according to a 1st Hermitian wavelet. The darkest part of the stimulus is located at the topmost point of the panel. Within the arena are marked tracks (297 in total), beginning near the centre, which delineate the paths taken by individual animals in trials. In these trials, animals must first negotiate an obstruction at the centre. Each colour corresponds to a given individual (of 50 in total). The points just after the track first crosses each of the two (dashed black and blue) concentric circles near the centre are used to mark a vector (shown with a straight line, dashed for headings directed to the target) to the arena wall. The intersection with the wall (the animal's heading) is shown by a hollow circle and the id of the animal. Animals are oriented in the 45d condition (and to a lesser extent the 60d and 150d condition) indicating that this species can see at a resolution better than 45d but worse than 30d.




#### Effect of differing sector sizes

```{r Plot success with differing sectors}
headings_summ |> pivot_longer(c(quart,quint,sext)) |>
  filter(obstacle=="obst") |>
  ggplot(aes(x=arc,y=value,color=name)) + geom_point() + geom_line() +
  geom_segment(x=0,xend=150,y=.2,yend=.2,color="darkgreen",linetype=2)+
  geom_segment(x=0,xend=150,y=1/6,yend=1/6,color="blue",linetype=2)+
  geom_segment(x=0,xend=150,y=1/4,yend=1/4,color="red",linetype=2)
```

Write the tracks to a file

```{r}
readr::write_csv(trackz_tbl,      file = "tracks_file.csv")
readr::write_delim(trackz_tbl,    file = "tracks_file.txt",delim = "\t")
readr::write_csv(trialz,  file = "headings_file.csv")
readr::write_delim(trialz,file = "headings_file.txt",delim = "\t")
```

### For Powerpoint Presentation

Get Hermitian(x_px) into a 0 to 1 bounded form and use that to make the grey scale.

The Hermitian formula won't accept a series without a sufficiently low x value. There needs to be a positive value of Gwv2(x), otherwise with x=0, 1/(max(x))= Inf. Need to make one for each arc.

```{r}
ptz      <- 3600
x_px     <- seq(from=-pi,to=pi,length.out=ptz)
arcz     <- sort(unique(df$arc))
S <- seq(0,2*pi,length.out=ptz)
S <- mapply(rad2clock,S)
S <- mapply(rad_shepherd_clock,S+pi)
Gwv       <- function(x){ (1 * exp(-x^2/(2*D^2))) }
Gwv2      <- function(x){ c(diff(Gwv(x)) / diff(x), 0)}
Gw_deriv  <- function(x){ Gwv2(x) * (1/max(Gwv2(x))) }
Hermitian <- function(x){(1 + Gw_deriv(x))/2}
df_Herm <- tibble( x=1.02*cos(S), y=1.02*sin(S)  )

for(i in 1:length(arcz)){
 T_px      <- (arcz[i]/360*pi)
 D         <- T_px / (2 * sqrt(2*log(2))) # cut out: sqrt(3) *
 z1 <- Hermitian(x_px)
 z1[is.na(z1) == TRUE] <- 0.5
 df_Herm[,i+2] <- scales::rescale(z1)
}

colnames(df_Herm) <- c("x","y",arcz)
df_Herm <- pivot_longer(df_Herm,cols = -c(x,y),names_to = "arc",values_to ="tone")
#df_Herm$arc <- as.factor(as.numeric(df_Herm$arc)) # make a factor but in numeric order
df_Herm$arc  <- as.numeric(df_Herm$arc) 

w_pattern_px <- 2*pi
plot(1, type="n", xlab="Angular position", ylab="Intensity", 
     xlim=c(0.5-w_pattern_px/2, -0.5+w_pattern_px/2), ylim=c(0, 1), xaxt='n')
     axis(1, seq(from=-w_pattern_px/2,to=w_pattern_px/2,length.out = 7),
             seq(from=-pi,to=pi,length.out = 7)  )
lines(x=x_px,y=Hermitian(x_px),col="green",lwd=2,lty=4)
```

Plot of relative tracks of obstructed animals for slide

```{r rel headings sheet fig, fig.height = 24.5/2.54, fig.width = 16/2.54}
library('ggnewscale')
obstruct  <- "obst"
linetypez <- c("FALSE"="dashed","TRUE"="solid")
trackz_tbl |>  filter(obstacle==obstruct) |> 
ggplot() + coord_fixed() + 
 facet_wrap(vars(arc),ncol=2,strip.position="top")+
geom_point(data=df_Herm,aes(x=x,y=y,color=tone,group=arc),size=1.5)+
  scale_color_gradient(low="black",high="white")+ new_scale("color")+  
 geom_path(aes(x=rel_x, y=rel_y, color=id))+ 
 annotate("path", color="blue",linetype="dashed",
 x=circ1*cos(seq(0,2*pi,0.05)),
 y=circ1*sin(seq(0,2*pi,0.05))) +
 annotate("path", color="black",linetype="dashed",
 x=circ0*cos(seq(0,2*pi,0.05)),y=circ0*sin(seq(0,2*pi,0.05))) +
 #annotate("text",x=c(1.1,0,-1.1,0),y=c(0,1.1,0,-1.1),label=c("90","0","270","180"))+
 #annotate("text", x=c(1.1,0),y=c(0,1.1),label=c("90","0")) +  
 annotate("text", x= .8, y= .8, label=paste0('45',"°"))  +  
 annotate("text", x= .8, y=-.8, label=paste0('135',"°")) +  
 annotate("text", x=-.8, y=-.8, label=paste0('225',"°")) +  
 annotate("text", x=-.8, y= .8, label=paste0('315',"°")) +
 geom_point(data=trialz[ # Stick on final relative headings
  trialz$obstacle==obstruct & as.numeric(trialz$id) <= how_many,],
  aes(x=cos(radians),y=sin(radians),color=id),size=2,shape=1)+scale_shape_identity()+
  #alpha=.5,scale_shape_manual(values="\u25CC")
 geom_segment(data=trialz[
   trialz$obstacle==obstruct & as.numeric(trialz$id) <= how_many,],
   aes(x=rel_x0,y=rel_y0,xend=cos(radians),yend=sin(radians),
     colour=id),linetype="dotted") +
 geom_text(data=headings_summ[headings_summ$obstacle %in% obstruct,], #panel labels
  aes(x=-.75,y=-1,label=paste('rho',"==",round(headings_summ$R[headings_summ$arc==arc],2))),
  size=4,parse=T)+ 
 geom_text(data=headings_summ[headings_summ$obstacle==obstruct,], #panel labels
  aes(x=-.9,y=1,label=paste0(arc,"°")),size=6,color="blue")+
 geom_segment(data=headings_summ[headings_summ$obstacle==obstruct,],aes(
  x=0,y=0,xend=R*cos(mu_r),yend=R*sin(mu_r),group=arc),
  size=1.3,colour="black",arrow=arrow(length=unit(.2,"cm")))+
 geom_segment(data=line_df,aes(x=x,y=y,xend=xend,yend=yend),inherit.aes=F) +
  xlab('') + ylab('') + theme_classic() + 
 geom_text_repel(data=trialz[trialz$obstacle==obstruct & as.numeric(trialz$id) <= how_many,],
   aes(x=cos(radians),y=sin(radians),color=id,label=id,max.overlaps=5),size=2.7)+#,check_overlap=T)+
 labs(title='')+#Plot tracks with obstacle relative to stimulus') +
 theme(axis.line=element_blank(),axis.ticks=element_blank(),
        axis.text=element_blank()) + theme(legend.position = "none") +
 theme(panel.spacing = unit(.5, "cm")) +
 theme(plot.title=element_text(size=14,hjust=0.5),#vjust=-62), 
        strip.background=element_blank(),strip.text=element_blank()) +
 theme(panel.background = element_rect(fill='transparent'), #transparent panel bg
   plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
   panel.grid.major = element_blank(), #remove major gridlines
   panel.grid.minor = element_blank(), #remove minor gridlines
   legend.background = element_rect(fill='transparent'), #transparent legend bg
   legend.box.background = element_rect(fill='transparent') #transparent legend panel
  )
ggsave('Plividus_prelim_rel_head.svg',bg='transparent')
```


```{r  rel headings sheet fig fun, fig.height = 6.625, fig.width = 10}
source('circ_ggplot_fun.R')
circ_ggplot_fun(trackz_tbl)
```



G1D_150_obstacle/WIN_20210118_15_39_44_Pro is weird - directly 'right' at 150d. Different video.

```{r}
trialz[trialz$arc==150 & trialz$radians > (pi/2)-.2 & trialz$radians < (pi/2)-0.1 & trialz$obstacle =="obst",]$video
```
This on was from the left. It could be the geom used to draw the lines. Not present in other plots.

It is a repeat from the same individual.

To make a Hermitian circle around the plot, make a colour profile using numbers for each degree. Make singleline function with only 360 numbers. Use scale_colour_grey()


### Unobstructed 150d powerpoint

```{r rel headings 150d unobst sheet fig, fig.height = 5.625, fig.width = 10}
library('ggnewscale')
#trackz_tbl$arc <- factor(trackz_tbl$arc)
df$id <- factor(df$id)
obstruct <- c("unobst","obst")#"unobst" #
arc_angs <- c(150)

linetypez <- c("FALSE"="dashed","TRUE"="solid")
trackz_tbl |>  filter(obstacle %in% obstruct & arc %in% arc_angs) |> 
ggplot() + coord_fixed() + 
 facet_wrap(vars(arc,obstacle),ncol=3,strip.position="top")+
 annotate("path",colour="black",x=cos(seq(0,2*pi,0.06)),y=sin(seq(0,2*pi,0.06)))+
  geom_point(data=df_Herm[df_Herm$arc %in% arc_angs,],
             aes(x=x,y=y,color=tone,group=arc),size=1.5)+
   scale_color_gradient(low="black",high="white")+
 new_scale("color")+  
 geom_path(aes(x=rel_x, y=rel_y,color=vid_ord,group=vid_ord))+ 
 annotate("path", color="blue",linetype="dashed",
  x=circ1*cos(seq(0,2*pi,0.05)),
  y=circ1*sin(seq(0,2*pi,0.05))) +
 annotate("path", color="black",linetype="dashed",
  x=circ0*cos(seq(0,2*pi,0.05)),y=circ0*sin(seq(0,2*pi,0.05))) +
 #annotate("text",x=c(1.1,0,-1.1,0),y=c(0,1.1,0,-1.1),label=c("90","0","270","180"))+
 #annotate("text", x=c(1.1,0),y=c(0,1.1),label=c("90","0")) +  
 annotate("text", x= .8, y= .8, label=paste0('45',"°"))  +  
 annotate("text", x= .8, y=-.8, label=paste0('135',"°")) +  
 annotate("text", x=-.8, y=-.8, label=paste0('225',"°")) +  
 annotate("text", x=-.8, y= .8, label=paste0('315',"°")) +
 geom_point(data=trialz[ # Stick on final relative headings
  trialz$obstacle==obstruct & trialz$arc %in% arc_angs & as.numeric(trialz$id) <= how_many,],
  aes(x=cos(radians),y=sin(radians),color=vid_ord),size=2,shape=1)+scale_shape_identity()+
 geom_text(data=headings_summ[
  headings_summ$obstacle %in% obstruct & headings_summ$arc %in% arc_angs,], #panel labels
  aes(x=-.75,y=-1,label=paste('rho',"==",round(headings_summ$R[headings_summ$arc==arc],2))),
  size=5,parse=T)+ 
 geom_segment(data=trialz[ trialz$obstacle %in% obstruct & 
     trialz$arc %in% arc_angs & as.numeric(trialz$id) <= how_many,],
   aes(x=rel_x0,y=rel_y0,xend=cos(radians),yend=sin(radians),
    colour=vid_ord),linetype="dotted") +
 geom_text(data=headings_summ[
  headings_summ$obstacle %in% obstruct & headings_summ$arc %in% arc_angs,], #panel labels
  aes(x=-.8,y=1,label=paste0(arc,"°")),size=6,color="blue")+
 geom_segment(data=headings_summ[
   headings_summ$obstacle %in% obstruct & headings_summ$arc %in% arc_angs,],
   aes(x=0,y=0,xend=R*cos(mu_r),yend=R*sin(mu_r),group=arc),
  size=1.3,colour="black",arrow=arrow(length=unit(.2,"cm")))+
 geom_segment(data=line_df,aes(x=x,y=y,xend=xend,yend=yend),inherit.aes=F) +
  xlab('') + ylab('') + theme_classic() + 
  labs(title='')+#Plot tracks with obstacle relative to stimulus') +
  theme(axis.line=element_blank(),axis.ticks=element_blank(),
        axis.text=element_blank()) + 
  theme(legend.position = "none") +
  theme(panel.spacing = unit(.5, "cm")) +
  theme(plot.title=element_text(size=14,hjust=0.5),#vjust=-62), 
        strip.background=element_blank(),strip.text=element_blank()) +
  theme(panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') #transparent legend panel
  )
ggsave('Plividus_prelim_rel_head_150d_unobst.svg',bg='transparent')
```

In future, geom_spoke() for angles.

Ultimately should make everything -90d because the top of the camera FOV actually pointed approx West.

Empty plot of relative tracks panels

```{r empty headings sheet fig, fig.height = 5.625, fig.width = 10}
library('ggnewscale')
#trackz_tbl$arc <- factor(trackz_tbl$arc)
obstruct  <- "obst"
linetypez <- c("FALSE"="dashed","TRUE"="solid")
trackz_tbl |>  filter(obstacle==obstruct & id==1) |> 
ggplot() + coord_fixed() + 
 facet_wrap(vars(arc),ncol=3,strip.position="top")+
 #annotate("path",colour="black",x=cos(seq(0,2*pi,0.06)),y=sin(seq(0,2*pi,0.06)))+
geom_point(data=df_Herm,aes(x=x,y=y,color=tone,group=arc),size=1.5)+
  scale_color_gradient(low="black",high="white")+
 new_scale("color")+  
 geom_path(aes(x=rel_x, y=rel_y, color=id))+ 
 annotate("path", color="blue",linetype="dashed",
 x=circ1*cos(seq(0,2*pi,0.05)),
 y=circ1*sin(seq(0,2*pi,0.05))) +
 annotate("path", color="black",linetype="dashed",
 x=circ0*cos(seq(0,2*pi,0.05)),y=circ0*sin(seq(0,2*pi,0.05))) +
 #annotate("text",x=c(1.1,0,-1.1,0),y=c(0,1.1,0,-1.1),label=c("90","0","270","180"))+
 #annotate("text", x=c(1.1,0),y=c(0,1.1),label=c("90","0")) +  
 annotate("text", x= .8, y= .8, label=paste0('45',"°"))  +  
 annotate("text", x= .8, y=-.8, label=paste0('135',"°")) +  
 annotate("text", x=-.8, y=-.8, label=paste0('225',"°")) +  
 annotate("text", x=-.8, y= .8, label=paste0('315',"°")) +
 geom_point(data=trialz[ # Stick on final relative headings
  trialz$obstacle==obstruct & as.numeric(trialz$id) <= 1,],
  aes(x=cos(radians),y=sin(radians),color=id),size=2,shape=1)+scale_shape_identity()+
  #alpha=.5,scale_shape_manual(values="\u25CC")
 geom_text(data=trialz[trialz$obstacle==obstruct & as.numeric(trialz$id) <= 1,],
   aes(x=.9*cos(radians),y=.9*sin(radians),color=id,label=id),size=2.7,check_overlap=T)+
 geom_segment(data=trialz[
   trialz$obstacle==obstruct & as.numeric(trialz$id) <= 1,],
   aes(x=rel_x0,y=rel_y0,xend=cos(radians),yend=sin(radians),
     colour=id),linetype="dotted") +
 geom_text(data=headings_summ[headings_summ$obstacle==obstruct,], #panel labels
  aes(x=-.9,y=1,label=paste0(arc,"°")),size=6,color="blue")+
 #geom_segment(data=headings_summ[headings_summ$obstacle==obstruct,],aes(
#  x=0,y=0,xend=R*cos(mu_r),yend=R*sin(mu_r),group=arc),
 # size=1.3,colour="black",arrow=arrow(length=unit(.2,"cm")))+
 geom_segment(data=line_df,aes(x=x,y=y,xend=xend,yend=yend),inherit.aes=F) +
  xlab('') + ylab('') + theme_classic() + 
  labs(title='')+#Plot tracks with obstacle relative to stimulus') +
  theme(axis.line=element_blank(),axis.ticks=element_blank(),
        axis.text=element_blank()) + theme(legend.position = "none") +
  theme(panel.spacing = unit(.5, "cm")) +
  theme(plot.title=element_text(size=14,hjust=0.5),#vjust=-62), 
        strip.background=element_blank(),strip.text=element_blank()) +
  theme(panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transparent legend bg
    legend.box.background = element_rect(fill='transparent') #transparent legend panel
  )
ggsave('Plividus_prelim_rel_head.svg',bg='transparent')
```


Empty plot of single panel

```{r empty 1 panel headings sheet fig, fig.height = 5.625, fig.width = 10}
library('ggnewscale') # for the gradients, allows multiple scales
#trackz_tbl$arc <- factor(trackz_tbl$arc)
obstruct  <- "obst"
arc_angs  <- c(45) 

linetypez <- c("FALSE"="dashed","TRUE"="solid")
trackz_tbl |>  filter(obstacle==obstruct & id==1 & arc %in% arc_angs) |> 
ggplot() + coord_fixed() + 
 #facet_wrap(vars(arc),ncol=3,strip.position="top")+
 #annotate("path",colour="black",x=cos(seq(0,2*pi,0.06)),y=sin(seq(0,2*pi,0.06)))+
geom_point(data=df_Herm[df_Herm$arc %in% arc_angs,],
  aes(x=x,y=y,color=tone,group=arc),size=1.5)+
  scale_color_gradient(low="black",high="white") + new_scale("color")+  
 geom_path(aes(x=rel_x, y=rel_y, color=id))+ 
 annotate("path", color="blue",linetype="dashed",
 x=circ1*cos(seq(0,2*pi,0.05)),
 y=circ1*sin(seq(0,2*pi,0.05))) +
 annotate("path", color="black",linetype="dashed",
 x=circ0*cos(seq(0,2*pi,0.05)),y=circ0*sin(seq(0,2*pi,0.05))) +
 #annotate("text",x=c(1.1,0,-1.1,0),y=c(0,1.1,0,-1.1),label=c("90","0","270","180"))+
 #annotate("text", x=c(1.1,0),y=c(0,1.1),label=c("90","0")) +  
 annotate("text", x= .8, y= .8, label=paste0('45',"°"))  +  
 annotate("text", x= .8, y=-.8, label=paste0('135',"°")) +  
 annotate("text", x=-.8, y=-.8, label=paste0('225',"°")) +  
 annotate("text", x=-.8, y= .8, label=paste0('315',"°")) +
 geom_point(data=trialz[ # Stick on final relative headings
  trialz$obstacle %in% obstruct & trialz$arc %in% arc_angs & as.numeric(trialz$id) <= 1,],
  aes(x=cos(radians),y=sin(radians),color=id),size=2,shape=1)+scale_shape_identity()+
  #alpha=.5,scale_shape_manual(values="\u25CC")
 geom_text(data=trialz[trialz$obstacle %in% obstruct & trialz$arc %in% arc_angs & as.numeric(trialz$id) <= 1,],
   aes(x=.9*cos(radians),y=.9*sin(radians),color=id,label=id),size=2.7,check_overlap=T)+
 geom_segment(data=trialz[
   trialz$obstacle==obstruct & trialz$arc %in% arc_angs  & as.numeric(trialz$id) <= 1,],
   aes(x=rel_x0,y=rel_y0,xend=cos(radians),yend=sin(radians),
     colour=id),linetype="dotted") +
 geom_text(data=headings_summ[headings_summ$obstacle %in% obstruct & headings_summ$arc %in% arc_angs,], #panel labels
  aes(x=-.9,y=1,label=paste0(arc,"°")),size=6,color="blue")+
 #geom_segment(data=headings_summ[headings_summ$obstacle==obstruct,],aes(
#  x=0,y=0,xend=R*cos(mu_r),yend=R*sin(mu_r),group=arc),
 # size=1.3,colour="black",arrow=arrow(length=unit(.2,"cm")))+
 geom_segment(data=line_df,aes(x=x,y=y,xend=xend,yend=yend),inherit.aes=F) +
  xlab('') + ylab('') + theme_classic() + 
  labs(title='')+#Plot tracks with obstacle relative to stimulus') +
  theme(axis.line=element_blank(),axis.ticks=element_blank(),
        axis.text=element_blank()) + theme(legend.position = "none") +
  theme(panel.spacing = unit(.5, "cm")) +
  theme(plot.title=element_text(size=14,hjust=0.5),#vjust=-62), 
        strip.background=element_blank(),strip.text=element_blank()) +
  theme(panel.background = element_rect(fill='transparent'), #transparent panel bg
    plot.background = element_rect(fill='transparent', color=NA), #transparent plot bg
    panel.grid.major = element_blank(), #remove major gridlines
    panel.grid.minor = element_blank(), #remove minor gridlines
    legend.background = element_rect(fill='transparent'), #transp legend bg
    legend.box.background = element_rect(fill='transparent') #transp leg panel
  )
ggsave('Plividus_prelim_rel_head.svg',bg='transparent')
```

The stimulus plotted above is far two wide.




#### Radian conversion

```{r}
## change top of arena to be at zero and rotation clockwise
# trialz$abs_rads <- conversion.circular(
#                             trialz$abs_rads, type="angles",
#                             units="radians",zero=pi/2,rotation="clock",
#                             template="none", modulo="asis")
# 
# trialz$radians <- conversion.circular(
#                             trialz$radians, type="angles",
#                             units="radians",zero=pi/2,rotation="clock",
#                             template="none", modulo="asis")
# 
# # make the radians aligned to the top and rotation to be clockwise
# trackz_tbl$abs_theta <- conversion.circular(
#                             trackz_tbl$abs_theta, type="angles",
#                             units="radians",zero=pi/2,rotation="clock",
#                             template="none", modulo="asis")
# 
# # make the radians aligned to the top and rotation to be clockwise
# trackz_tbl$rel_theta <- conversion.circular(
#                             trackz_tbl$rel_theta, type="angles",
#                             units="radians",zero=pi/2,rotation="clock",
#                             template="none", modulo="asis")
```



